### 2013-10-01 doc/ms1.tex

\documentclass{emulateapj}

\usepackage{apjfonts}
\usepackage{epsfig}
\usepackage{natbib}

\newlength{\colwidth}\setlength{\colwidth}{3.5in}

\addtolength\voffset{0.8cm}

\newcommand{\ds}{\displaystyle}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\bee}{\begin{eqnarray}}
\newcommand{\eee}{\end{eqnarray}}

\newcommand{\ov}{ \Omega_{\rm \Lambda} }
\newcommand{\om}{ \Omega_{\rm m} }
\newcommand{\mob}{ M_{\rm obj} }
\newcommand{\mstar}{ M_{*}(a) }
\newcommand{\msol}{\hbox{${\rm M}_\odot$}}

\newcommand{\rvir}{\hbox{$r_{\rm 200}$}}
\newcommand{\mvir}{\hbox{$M_{\rm vir}$}}
\newcommand{\Mvir}{\hbox{$M_{\rm vir}$}}
\newcommand{\vvir}{\hbox{$v_{\rm 200}$}}
\newcommand{\tvir}{T_{\rm vir}}
\newcommand{\mthresh}{M_{\rm t}}
\newcommand{\nsats}{n_{\rm sats}}
\newcommand{\vmax}{{\rm v}_{\rm max}}
\newcommand{\zreion}{z_{\rm reion}}
\newcommand{\LCDM}{$\Lambda$CDM }
\newcommand{\sigate}{ \sigma_{8}}
\newcommand{\actoy}{a_{c,{\rm toy}}}
\newcommand{\K}{{\rm K}}

\newcommand{\DME}{D_{\rm M31}}
\newcommand{\vMEI}{v_{\rm M31}}
\newcommand{\MMEI}{M_{\rm M31}}
\newcommand{\DMEE}{D_{\rm M33}}
\newcommand{\vMEE}{v_{\rm M33}}
\newcommand{\MMEE}{M_{\rm M33}}
\newcommand{\MMW}{{\rm M}_{\rm MW}}
\newcommand{\Nsubs}{N_{\rm subs}}

\def\rx{X}
\def\ry{Y}
\def\rz{Z}
\def\vx{v_X}
\def\vy{v_Y}
\def\vz{v_Z}
\def\distance{D}
\def\vrad{v_{\rm rad}}
\def\vtan{v_{\rm tan}}

\def\pr{{\rm Pr}}
\def\pars{\mathbf{x}}
\def\parsi{x_i}
\def\data{\mathbf{d}}
\def\datai{d_i}
\def\datap{\mathbf{d}^{\rm p}(\pars)}
\def\datapi{d^{\rm p}_{i}(\pars)}

\newcommand{\Msun}{{\rm M}_\odot}
\newcommand{\hinv}{h^{-1}}
\newcommand{\Mpc}{{\rm Mpc}}
\newcommand{\hmpc}{\hinv\Mpc}
\newcommand{\kpc}{{\rm kpc}}
\newcommand{\kms}{{\rm km s}^{-1}}

\newcommand{\rhobar}{\langle \rho \rangle}
\newcommand{\sig}{ \langle \sigma v \rangle }
\newcommand{\sigwig}{ {\widetilde \sigma} }
\newcommand{\rbh}{ r_{\rm bh} }
\newcommand{\rmax}{ r_{\rm max} }
\newcommand{\kmsmpc}{\, \rm{km}\,  \rm{s}^{-1}\, \rm{Mpc}^{-1}}
\newcommand{\lsim}{\lower.5ex\hbox{\ltsima}}
\newcommand{\gsim}{\lower.5ex\hbox{\gtsima}}
\newcommand{\ltsima}{$\; \buildrel < \over \sim \;$}
\newcommand{\gtsima}{$\; \buildrel > \over \sim \;$}
\def \etal      {\hbox{et al.} }

\newcommand{\bolshoi}{{\sc Bolshoi }}
\newcommand{\consuelo}{{\sc Consuelo }}
\newcommand{\subfind}{\texttt{\footnotesize SUBFIND }}
\newcommand{\CMBFAST}{\texttt{\footnotesize CMBFAST }}

\def\Sref#1{Section~\ref{#1}}
\def\Fref#1{Figure~\ref{#1}}
\def\Tref#1{Table~\ref{#1}}
\def\Eref#1{Equation~\ref{#1}}
\def\Aref#1{Appendix~\ref{#1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% RESULTS:

\def\MLG{M_{\rm LG}}
\def\MMW{M_{\rm MW}}
\def\MEI{M_{\rm M31}}
\def\MEE{M_{\rm M33}}

% Local group mass: MW+M31 from M31 kinematics:
\def\MPAIRestimate{XX}
\def\MPAIRerrorplus{xx}
\def\MPAIRerrorminus{xx}

% Local group timing argument quantities:
\def\ATA{A_{200}}
\def\MTA{M_{LG,TA}}

% Mass of M31+M33 system, from M33 kinematics:
\def\MbAestimate{XX}
\def\MbAerrorplus{xx}
\def\MbAerrorminus{xx}

% Mass of local group, MW+M31, from both M31 and M33 kinematics:
\def\MTRIPLETestimate{XX}
\def\MTRIPLETerrorplus{xx}
\def\MTRIPLETerrorminus{xxx}

% Individual galaxyhalo mass estimates from both M31 and M33 kinematics:
\def\MEIestimate{XX}
\def\MEIerrorplus{xx}
\def\MEIerrorminus{xx}

\def\MEEestimate{XX}
\def\MEEerrorplus{xx}
\def\MEEerrorminus{xx}

\def\MMWestimate{XX}
\def\MMWerrorplus{xx}
\def\MMWerrorminus{xx}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage[usenames]{color}
\newcommand{\risa}[1]{\textcolor{red}{\bf #1}}
\newcommand{\phil}[1]{\textcolor{ForestGreen}{\bf #1}}
\newcommand{\mike}[1]{\textcolor{Violet}{\bf #1}}
\newcommand{\question}[1]{\textcolor{blue}{\bf #1}}
\newcommand{\todo}[2]{{\bf To do (#1): #2}}
\newcommand{\new}[1]{{\bf #1}}

\newenvironment{inlinefigure}{\def\@captype{figure}

\noindent\begin{minipage}{0.999\linewidth}\begin{center}}
{\end{center}\end{minipage}\smallskip}

\citestyle{aa}
\shortauthors{Marshall et al.}
\shorttitle{The Mass of the Local Group}

% ----------------------------------------------------------------------------

\begin{document}

\title{The Dark Matter Distribution of the Local Group inferred from
Cosmological Simulations: I. The Halo Masses of M31, M33 and the Milky Way}

\author{Philip J. Marshall\altaffilmark{1}}
\author{Michael T. Busha\altaffilmark{2}} 
\author{Risa H. Wechsler\altaffilmark{3, 4}}
\author{Peter Behroozi\altaffilmark{3}}
\author{XX\altaffilmark{5}}
\author{YY\altaffilmark{6}}
\altaffiltext{1}{Department of Physics, Oxford University, Oxford, OX1 3RH, UK.} 
\altaffiltext{2}{Institute for Theoretical Physics, University of
 Zurich, 8057 Zurich, Switzerland}  
\altaffiltext{3}{Kavli Institute for Particle Astrophysics and Cosmology,
Physics Department, Stanford University, Stanford, CA, 94305, USA
{\tt mbusha@physik.uzh.ch, pjm@slac.stanford.edu, rwechsler@stanford.edu}}
\altaffiltext{4}{Particle and Particle Astrophysics Department, 
SLAC National Accelerator Laboratory, Menlo Park, CA, 94025, USA}
\altaffiltext{5}{XX}
\altaffiltext{6}{YY}

% ----------------------------------------------------------------------------

\begin{abstract}

We study the local group of galaxies in its cosmological context, by matching
dark matter halos in the \consuelo N-body simulation with kinematic
observations of M31 and M33. Treating the \consuelo halo  catalog as a list of
samples drawn from the cosmological prior PDF for the local group properties,
we select local group analogs whose isolation matches our own, and then
compute importances for the samples as combined likelihoods for the
observations. 
%
From the ``observed'' galactocentric distance, radial velocity and tangential 
velocity of M31 alone, we infer the mass of the local group to be $\MLG =
(\MMW + \MEI) = (\MPAIRestimate^{+\MPAIRerrorplus}_{-\MPAIRerrorminus}) \times
10^{12} \Msun$. This is ... than the mass that has been estimated from the
timing argument: we show that for our matched analogs the timing argument mass
systematically ... the local group mass. 
%
Transforming the ``observed'' galactocentric distance, radial velocity and
tangential  velocity of M33 into the M31 rest-frame, and accounting for
covariance between these measurements, we also infer the mass of the M31--M33
system to be  $(\MEI + \MEE) =
(\MbAestimate^{+\MbAerrorplus}_{-\MbAerrorminus}) \times 10^{12} \Msun$.
%
Combining the kinematic constraints from both M31 and M33,  we infer a local
group mass of $\MLG = (\MMW + \MEI) =
(\MTRIPLETestimate^{+\MTRIPLETerrorplus}_{-\MTRIPLETerrorminus}) \times
10^{12} \Msun$, and are able to estimate the halo masses of the three galaxies
individually, finding these to be $\MEI =
(\MEIestimate^{+\MEIerrorplus}_{-\MEIerrorminus}) \times 10^{12} \Msun$, $\MEE
= (\MEEestimate^{+\MEEerrorplus}_{-\MEEerrorminus}) \times 10^{12} \Msun$, 
and $\MMW = (\MMWestimate^{+\MMWerrorplus}_{-\MMWerrorminus}) \times 10^{12}
\Msun$. 
%
This Milky Way halo mass estimate is in xxxagreement with previous estimates
from stellar and dwarf satellite kinematics.

\end{abstract}

\keywords{%
Galaxy: halo, fundamental parameters -- 
galaxies: haloes, fundamental parameters -- 
galaxies: Local Group -- 
galaxies: individual: M31, M33 --
cosmology: dark matter}

% ----------------------------------------------------------------------------

\section{Introduction}
\label{sec:intro}

The Local Group (LG) of galaxies consists of two large spiral galaxies,
Andromeda (M31) and the Milky Way (MW), and a population of smaller
satellite galaxies, the biggest of which is Triangulum (M33). We know
relatively little about the distribution of dark matter throughout the Local
Group volume, beyond simple estimates of the total mass of the group
estimated from its members kinematics \citep[e.g.\ ][hereafter VG08]{vdM+G08}.
Our knowledge of the dark matter halos of the group member galaxies is
similarly sparse: virial masses have been estimated from extrapolations of
simple models fitted to stellar or gas kinematics \citep[e.g.\
][]{Xue+08,M31rotcurve} or to the observed positions and 
velocities of satellite galaxies \citep[e.g.\ ][]{EvansEtc}.

Our goal is to map out the dark matter distribution in the Local Group
by combining observational data on the  positions and velocities of the
three largest LG members with our understanding of structure formation
in $\Lambda$CDM universes, as detailed in halo catalogs extracted from
cosmological N-body simulations. We treat halo catalogs as containing
samples drawn from a very informative, highly correlated prior PDF for
halo properties, and then  importance sample them with likelihood
functions based on the observed kinematics of the particular system
under study. In \citet[][hereafter B11]{Bus++11} we presented a new
measurement of the Milky Way halo mass, made in this way using the
\bolshoi simulation of \citet{Bolshoi} and the distances to and speeds
of the Magellanic Clouds. In this work, we use the lower resolution but
higher volume \consuelo simulations, and constrain them with the radial
velocities, proper motions and distances of M31 and M33, as well as our
previous MW mass estimate.

The dynamics of the Local Group system, captured by the constrained
\consuelo halos, are the same as feature in the Timing Argument of
\citet{K+W59}; the difference is that instead of a point mass on a
radial orbit, our model for the motion of M31 is only that it was
generated in a $\Lambda$CDM universe. We can, however, follow
\citet{L+W08} and ``calibrate'' the Timing Argument, effectively asking
what observers in MW-like systems would infer about their LG-like group
mass based on the Timing Argument.  Meanwhile, the masses of the LG and
each of its three most massive constituents are accessible directly once
we have assumed the \consuelo catalogue to contain a representative
sample of the dark matter halos found in our universe. In a subsequent
paper we will explore other properties of these halos.

In short, we seek answers to the following questions:
\begin{itemize}
\item What is the mass of dark matter in the Local Group?
\item How accurate is the Timing Argument for the Local Group mass?
\item What are the masses of the dark matter halos of M31, M33 and the Milky
Way?
\end{itemize}

This paper is structured as follows. In \Sref{sec:method} we review the
halo inference methodology, and introduce the \consuelo N-body
simulation suite we use. Then in \Sref{sec:data} we summarise the
observational constraints we place on our model halo population during
the importance sampling process: this includes the isolation criteria we
use in the initial halo selection step. In \Sref{Sec:results} we present
our results on the mass of the Local Group and its members, and then
discuss those results in the context of previous work in
\Sref{sec:discuss}. We conclude in \Sref{sec:conclude}.

Throughout this work we assume a \LCDM cosmology, with $\Omega_{\rm m} =
0.3$, $\Omega_{\rm \Lambda} = 0.7$, and $H_0 = 70 \kms \Mpc^{-1}$. Where
a point estimate of a parameter is given, its value is the median of the
one-dimensional marginalised posterior probabilty distribution, and its
quoted uncertainty describes the 68\% credible region.
We define a halo's mass by its $M_{200,\rm c}$, the mass enclosed within a
sphere that is over-dense by a factor of 200 relative to the critical density
at the redshift under consideration.

% ----------------------------------------------------------------------------

\section{Methodology}
\label{sec:method}

In this section we review the halo inference introduced in B11, and describe
the \consuelo cosmological N-body simulations that we use in this work.

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\subsection{Halo catalogs: samples from the halo prior PDF}
\label{sec:sampling}

\todo{Phil}{Write this section!}

Basic concept. Multivariate PDF fof halo properties.

Importance sampling by likelihood of observed properties.

Protection against over-constraint: measures of effective sample number,
bootstrapping.

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\subsection{The \consuelo Simulation}
\label{sec:sampling}

\todo{Mike}{Write this section!}

Description: size, resolution. Number of halos. Mvir and cvir converted to M200
and c200?

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\subsection{The Local Group prior}
\label{sec:sampling}

\todo{Mike}{Write this section!}

Justification of hard-edged prior. Isolation criteria. Comparison to work of
CLUES (constrained sims), Li \& White (timing arg), and Karenchentsev (local
void).

Definition of pair, triplet (subset of pairs). Resulting numbers.

Figure showing MW, M31, M33, MLG mass prior for pairs and triplets. Comment:
triplets have more mass than pairs, how big an effect?

% ----------------------------------------------------------------------------

\section{Observational Constraints}
\label{sec:data}

\todo{Phil}{Write this section!}

We now turn to the observational data that we use to constrain our two and
three-halo cosmological models of the Local Group. Separate out
the constraints by galaxy, M31 and M33.  Explain approach of shifting to M31
frame. Orbit phase not interesting; M33 motion relative to MW not very
interesting.

Figure: M31/MW Pr(D, vr, vt) with and without constraints on all these. Show
impact of having an M33.

Figure: M33 Pr(D, vr, vt) with and without constraints on all these. All M33-M31
systems have an MW.

Comment on number of samples remaining in both pairs and triplets cases. Which
observed properties are unusual, if any?

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\def\arraystretch{1.4}
\begin{deluxetable}{cccc}[ht!]
\tabletypesize{\footnotesize}
\tablecaption{\label{table:properties}
Observed kinematic properties of M31 and M33.}
\startdata
\hline
\hline
                      & M31               & M33               & Reference          \\
\hline
$\rx$ / kpc           & $-378.9 \pm 30.?$ & $-476.1 \pm 30.?$ & vdM12   \\
$\ry$ / kpc           & $ 612.7 \pm 30.?$ & $ 491.1 \pm 30.?$ & vdM12   \\
$\rz$ / kpc           & $-283.1 \pm 30.?$ & $-412.9 \pm 30.?$ & vdM12   \\
$\vx$ / km s$^{-1}$   & $  66.1 \pm 26.7$ & $  43.1 \pm 21.3$ & vdM12   \\
$\vx$ / km s$^{-1}$   & $ -76.3 \pm 19.0$ & $ 101.3 \pm 23.5$ & vdM12   \\
$\vx$ / km s$^{-1}$   & $  45.1 \pm 26.5$ & $ 138.8 \pm 28.1$ & vdM12   \\
\hline
                      & MW                & M33               & \\
\hline
$\distance$ / kpc     & $ xxx.x \pm xx.x$ & $ xxx.x \pm xx.x$ & \\
$\vrad$ / km s$^{-1}$ & $  xx.x \pm xx.x$ & $  xx.x \pm xx.x$ & \\
$\vtan$ / km s$^{-1}$ & $  xx.x \pm xx.x$ & $  xx.x \pm xx.x$ & \\
\hline
\enddata
\tablecomments{
$(\rx,\ry,\rz)$ is a position vector relative to the Galactic center, while
$(\vx,\vy,\vz)$ is a three-dimensional Galactocentric velocity vector. The
components of these vectors were used to derive distances and  radial and
tangential velocities {\it relative to M31}: these are the inputs to the
likelihood calculations in the text. Uncertainties on quantities 
in the lower part of this
table were calculated by Monte
Carlo, assuming uncorrelated errors on quantities in the upper part.
%
References are:
vdM02~=~\citet{vanderMarel12}
}
\medskip
\end{deluxetable}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% ----------------------------------------------------------------------------

\section{Results}
\label{sec:results}

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\subsection{The Mass of the Local Group from the Kinematics of M31}
\label{sec:results:LGmass}

Pairs: D, vr, vt constraints from M31. 
Measurement of  MLG. Precision.

Figure: 1D plot of Pr(MLG) given \\
  1) M31 distance\\
  2) M31 vr\\
  3) M31 vt\\
  4) M31 all kinematics

Corner plot of Pr(MMW, MM31, MLG), prior vs all constraints

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\subsection{The Mass of M31 from the Kinematics of M33}
\label{sec:results:M31mass}

Couples: D, vr, vt constraints from M33 given an M31-MW pair. 
Measurement of  MM31. Precision.

Figure: 1D plot of Pr(MM31) given \\
  1) M33 distance\\
  2) M33 vr\\
  3) M33 vt\\
  4) M33 all kinematics

Corner plot of Pr(MM31, MM33, Mcouple), prior vs all constraints

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\subsection{The Masses of the Local Group Galaxy Halos from the Kinematics of M31 and M33}
\label{sec:results:jointanalysis}

Triplets: D, vr, vt constraints from M33 and M31. 
Measurement of MMW. Precision.

Figure: 1D plot of Pr(MMW) given \\
  1) M31 kinematics\\
  2) + M33 kinematics\\

Figure: Corner plot of Pr(MMW, MM31, MM33, MLG) given\\
   1) prior \\
   2) all kinematics \\

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\subsection{Comparison with the Local Group Timing Argument Mass Estimate}
\label{sec:results:TA}

Timing argument introduction. Review of Li \& White. Two halo sum, M31
kinematics, "radial" MTA. Review of VdM12's comment on precision - aim is to
update Li\&White, show improvement in precision.

Can compute radial MTA for every model local group,  and compare with actual
sums of M200. "Calibration" in sense that MTA gives estimate of MW + M31
masses, under assumption they are point masses on Keplerian orbits. Bias and
scatter both important.

Figure: corner plot of MLG, MTA, A200 given \\
  1) M31 kinematics (pairs+DvrM31)\\
  2) M31 + M33 kinematics (triplets+DVrM31+DVrM33)\\

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\subsection{Focusing on the Milky Way: constraints from other M31 and M33 halo mass estimates}
\label{sec:results:MW}

Review other measurements of M31 and M33 halos. Adopt one, possibly from
VdM12. Plot impact on Pr(MMW, MM31, MM33, MLG) and MW mass estimate. MW mass
estimate, precision. Tension?

Figure: Corner plot of Pr(MMW, MM31, MM33, MLG) given\\
   1) M31 and M33 kinematics (from earlier) \\
   2) these plus M31 and M33 halo mass estimates. \\

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

%%%%%%%%%%%%%%%%%%%%%%
% \input{mass_table.tex}
%%%%%%%%%%%%%%%%%%%%%%

% ----------------------------------------------------------------------------

\section{Discussion}
\label{sec:discuss}

% \subsection{Statistical Robustness}
% \label{sec:bootstrap}
% 
% \subsection{The Information Content of the Observations}
% \label{sec:bootstrap}
% Improvements in measurements from cosmological prior? eg distances?
% 
% 
% \subsection{The Masses of M31 and M33: Comparison with Stellar Dynamics
% Results}
% \label{sec:M31M33stellardynamics}
% 
% 
% \subsection{The Local Group Mass: Comparison with the Timing Argument}
% \label{sec:timing}
% 
% 
% \subsection{Further Work}
% \label{sec:future}
% 
% ----------------------------------------------------------------------------

\section{Conclusions}
\label{sec:conclude}

From this analysis we draw the following conclusions:
\begin{itemize}

\item The \consuelo simulation has sufficient volume to permit analogs of
individual galaxy groups to be identified in large numbers. We identified XX
3-member groups with similar isolation properties to the Local Group.

\item Importance-sampling the \consuelo-sampled halo parameter prior PDF with
the probability of the M31 and M33 distances and velocities, and the mass of
the Milky Way, we infer a two-halo Local Group mass of $\MLG = (\MMW + \MEI)
= (\MPAIRestimate^{+\MPAIRerrorplus}_{-\MPAIRerrorminus}) \times 10^{12}
\Msun$.

\item M31-MW timing mass argument superceded. How accurate was it?

\item Weighed M31 using M33 kinematic information. Result.

\item Combining M31 and M33 kinematics to weigh halos. Inferences of the three
halo masses. MW mass. Tension with other datasets?

% \item We find that the M33 halo provides a significant addition to the Local
% Group mass: $\MEE = (\MEEestimate^{+\MEEerrorplus}_{-\MEEerrorminus}) \times
% 10^{12} \Msun$. Including the M33 halo increases the inferred Local Group mass
% to $\MTRIPLET = (\MMW + \MEI + \MEE) =
% (\MTRIPLETestimate^{+\MTRIPLETerrorplus}_{-\MTRIPLETerrorminus}) \times
% 10^{12} \Msun$. We infer the mass of M31 to be $\MEI =
% (\MEIestimate^{+\MEIerrorplus}_{-\MEIerrorminus}) \times 10^{12} \Msun$
% 
% \item For each sample Local Group analog, we can compute the timing argument
% mass that an observer at the Earth analog's position would derive: the PDF for
% this quantity is ... relative to that for the true Local Group mass. We find
% that ... 
% 
% \item The M33 kinematics contain information about the proper motion of M31:
% we predict this to be...

\end{itemize}

% ----------------------------------------------------------------------------

\bibliographystyle{apj}
\bibliography{references}

% ----------------------------------------------------------------------------

\end{document}

### 2013-10-02 LGHPinfer.py
Pseudo-code following discussion with MB

@@ -355,7 +355,7 @@ def LGHPinfer(argv):
   # Read in constraints as an instance of the constraints class, which 
   # contains the relevant observational data:
     
-  obs = LGconstraints.bundle()
+  obs = localgroup.bundle()
   
   # Calculate importances based on input MW and M31 constraints.
   

### 2013-10-02 localgroup/constraints.py

@@ -21,7 +21,6 @@ class bundle:
 # The actual observational data, for named parameters: 
    def assignment(self):
 
-
       # Version 1, radial velocity + distance relative to MW:
       
       # Milky Way:
@@ -35,30 +34,6 @@ class bundle:
       self.data['M33']['D']  = [0.799,0.023,gaussian] # VG08, Mpc
       self.data['M33']['vr'] = [-190.0,59.0,gaussian] # Brunthaler, km/s
       
-      # self.data['M33']['angsep'] = [14,8,0.1,gaussian] # ? Angular separation
-      
-      
-#       # Version 2, radial velocity + distance, relative to M31:
-# 
-#       # MW, relative to M31:
-#       # NEED TO UPDATE TO VdM++2012!
-#       self.data['MW']['D']  = [0.774,0.040,gaussian]  # VG08, Mpc
-#       self.data['MW']['vr'] = [-130.0,8.0,gaussian]   # VG08, km/s
-#       # self.data['MW']['vt'] = [0.0,4.0,halfgaussian]  # VG12, km/s
-# 
-#       # M33, relative to M31:
-#       # NEED TO COMPUTE FROM VdM++2012!
-#       self.data['M33']['D']  = [0.799,0.023,gaussian] # VG08, Mpc
-#       self.data['M33']['vr'] = [-190.0,59.0,gaussian] # Brunthaler, km/s
-#       # self.data['M33']['vt'] = [0.0,4.0,halfgaussian]  # To be calculated!!
-# 
-#       # Optional extras - component masses:
-# 
-#       # Milky Way mass:
-#       self.data['MW']['Mvir'] = [12.2,0.1,lognormal] # Busha++2011
-#       
-#       # M31 mass:
-#       self.data['M31']['Mvir'] = [12.4,0.1,lognormal] # ???
 
       return
 
@@ -88,3 +63,27 @@ def lognormal(x,x0,sigma):
    return gaussian(y,x0,sigma)   
 
 # ======================================================================
+      # self.data['M33']['angsep'] = [14,8,0.1,gaussian] # ? Angular separation
+      
+      
+#       # Version 2, radial velocity + distance, relative to M31:
+# 
+#       # MW, relative to M31:
+#       # NEED TO UPDATE TO VdM++2012!
+#       self.data['MW']['D']  = [0.774,0.040,gaussian]  # VG08, Mpc
[…]

### 2013-10-02 localgroup/halo.py

@@ -1,0 +1,104 @@
+# ======================================================================
+
+import localgroup
+
+import numpy
+
+# ======================================================================
+
+class Halo(object):
+    """
+    NAME
+        Halo
+
+    PURPOSE
+        Define a dark matter halo (or set of halos), that has either:
+         a) observables, like sky position, proper motion etc, or
+         b) cartesian phase space coordinates, {x,y,z,vx,vy,vz}
+        and potentially
+         c) properties, like Mass, Angular Momentum, etc,
+        and has the means to translate between its coordinate systems, 
+        transform observables to coordinates, and compute its likelihood
+        of being a local group analog.
+
+    COMMENTS
+        This class is used in two places: derivation of the local group
+        analog likelihood function, and then inference of local group
+        halo properties using that likelihood.
+
+    INITIALISATION
+
+    
+    METHODS
+
+        read
+        sample_from
+        translate_to
+        transform_to
+
+    BUGS
+
+    AUTHORS
+      This file is part of the LocalGroupHaloProps project, 
+      distributed under the GPL v2, 
+      by Phil Marshall and Michael Busha (KIPAC). 
+      Please cite: Marshall et al in preparation.
+
+    HISTORY
+      2013-10-01  started Marshall & Busha (KIPAC)
+    """
+# ======================================================================
+
+    def __init__(self):
+    
+        return
+        
+# ----------------------------------------------------------------------------
+
+    def read(self,halofile):
+        
+        self.properties = astropy.Table(halofile)
+        
+        self.x = self.properties['x']
+        self.y = self.properties['y']
+        self.z = self.properties['z']
+        self.vx = self.properties['vx']
+        self.vy = self.properties['vy']
+        self.vz = self.properties['vz']
+
+        return
+        
+# ----------------------------------------------------------------------------
+
+    def sample_from(self,obs,Nsamples):
+        
+        self.ra = draw(obs['RA'],Nsamples)
+        # etc
+        
+        return
+        
+# ----------------------------------------------------------------------------
+
+    def translate_to(self,other):
+        
+        self.x -= other.x
+        self.y -= other.y
+        self.z -= other.z
+        self.vx -= other.vx
+        self.vy -= other.vy
+        self.vz -= other.vz
+        self.frame = other.name
+        
+        return
+        
+# ----------------------------------------------------------------------------
+
+    def transform_to(self,newsystem):
+        
+        if newsystem == 'Heliocentric Cartesians' then:
+            self.coordinate_system == newsystem
+            pass
[…]

### 2013-10-02 localgroup/observations.py

@@ -1,0 +1,62 @@
+# ======================================================================
+
+import localgroup
+
+import numpy
+
+# ======================================================================
+
+class Observations(object):
+    """
+    NAME
+        Observations
+
+    PURPOSE
+        Encode all the observational information we have about the 
+        kinematics of the 3 largest local group galaxies.
+
+    COMMENTS
+
+    INITIALISATION
+
+    METHODS
+
+    BUGS
+
+    AUTHORS
+      This file is part of the LocalGroupHaloProps project, 
+      distributed under the GPL v2, 
+      by Phil Marshall and Michael Busha (KIPAC). 
+      Please cite: Marshall et al in preparation.
+
+    HISTORY
+      2013-10-01  started Marshall & Busha (KIPAC)
+    """
+# ======================================================================
+
+    def __init__(self):
+        self.data = dict()
+        self.data['MW'] = dict()
+        self.data['M31'] = dict()
+        self.data['M33'] = dict()
+        self.assignment()
+        return 
+
+# ----------------------------------------------------------------------------
+ 
+    def assignment(self):
+      
+        # Milky Way:
+        self.data['MW']['RA'] = [value,err,func]
+
+        # M31:
+        self.data['M31']['D']  = [0.774,0.040,'Gaussian'] # VG08, Mpc
+        self.data['M31']['vr'] = [-130.0,8.0,'Gaussian']  # VG08, km/s
+
+        # M33:
+        self.data['M33']['D']  = [0.799,0.023,'Gaussian'] # VG08, Mpc
+        self.data['M33']['vr'] = [-190.0,59.0,'Gaussian'] # Brunthaler, km/s
+
+        return
+
+# ======================================================================

### 2013-10-02 localgroup/system.py

@@ -1,0 +1,100 @@
+# ======================================================================
+
+import localgroup
+
+import numpy
+
+# ======================================================================
+
+class System(object):
+    """
+    NAME
+        System
+
+    PURPOSE
+        Define a local group analog, containing three halos, and 
+        perform various calculations.
+
+    COMMENTS
+        This class is used in two places: derivation of the local group
+        analog likelihood function, and then inference of local group
+        halo properties using that likelihood.
+
+    INITIALISATION
+
+    
+    METHODS
+
+        read_halos
+        observe_halos
+        get_parameters
+        get_properties
+
+    BUGS
+
+    AUTHORS
+      This file is part of the LocalGroupHaloProps project, 
+      distributed under the GPL v2, 
+      by Phil Marshall and Michael Busha (KIPAC). 
+      Please cite: Marshall et al in preparation.
+
+    HISTORY
+      2013-10-01  started Marshall & Busha (KIPAC)
+    """
+# ======================================================================
+
+    def __init__(self):
+    
+        return
+        
+# ----------------------------------------------------------------------------
+
+    def read_halos(self,directory):
+        
+        # Could be a lot more elegant.
+        
+        MW = localgroup.Halo()
+        MWfile = directory+'/box01_MW.fits'
+        MW.read(MWfile)
+        
+        M31 = localgroup.Halo()
+        M31file = directory+'/box01_M31.fits'
+        M31.read(MWfile)
+        
+        M33 = localgroup.Halo()
+        M33file = directory+'/box01_M33.fits'
+        M33.read(MWfile)
+        
+        return
+        
+# ----------------------------------------------------------------------------
+
+    def observe_halos(self):
+        
+        obs = localgroup.Observations()
+        Nsamples = 1000
+        
+        MW = localgroup.Halo()
+        MW.sample_from(obs.data['MW'],Nsamples)
+        
+        M31 = localgroup.Halo()
+        M31.sample_from(obs.data['M31'],Nsamples)
+        
+        M33 = localgroup.Halo()
+        M33.sample_from(obs.data['M33'],Nsamples)
+        
+        return
+        
+# ----------------------------------------------------------------------------
+
+    def get_parameters(self):
+        
+        return
+        
+# ----------------------------------------------------------------------------
+
+    def get_properties(self):
+        
+        return
+        
+# ======================================================================

### 2014-12-09 localgroup/likelihood.py
wrote evaluate method.

@@ -80,6 +80,15 @@ class Likelihood(object):
             raise ValueError("Unrecognised approximation mode %s" % mode)
 
         return
+# ======================================================================
+
+    def evaluate(self, points, mode="GMM"):
+        if (mode == "GMM"):
+            logprobs, resps = self.PDF.score_samples(points)
+        else:
+            raise ValueError("Unrecognised approximation mode %s" % mode )
+
+        return logprobs
 
 # ======================================================================
 
@@ -109,8 +118,8 @@ class Likelihood(object):
         train_num = np.rint(0.7*self.samples.shape[0])
         train_data = self.samples[:train_num]
         test_data = self.samples[train_num:]
-        for i in range(1,11):
-            pdf = mixture.GMM(i)
+        for i in range(1,maxMM+1):
+            pdf = mixture.GMM(i, covariance_type='full')
             pdf.fit(train_data)
             aic_scores.append(pdf.aic(test_data))
             bic_scores.append(pdf.bic(test_data))
@@ -120,8 +129,8 @@ class Likelihood(object):
 
 # ======================================================================
     def gaussianOverlay(self, figure):
-        n_gaussians = 2
-        colors = ['g', 'r', 'y']
+        n_gaussians = 5
+        colors = ['g', 'r', 'y', 'b', 'c']
         transparency = 0.5
         model = mixture.GMM(n_gaussians, covariance_type='full')
         axes = np.reshape(figure.axes, (6,6))

### 2014-12-09 localgroup/likelihood.py
added preprocessing of L.samples

@@ -24,11 +24,13 @@ class Likelihood(object):
 
     
     METHODS
-        generate         - draw sample vector from observations' distribuions
-        approximate      - compute KNN/GMM/etc estimate of PDF
-        evaluate         - compute value of PDF at given vector
-        plot_samples     - make triangle plot for observed data in M31 ref frame
-        set_PDF          - set the L.PDF field
+        generate           - draw sample vector from observations' distribuions
+        approximate        - compute KNN/GMM/etc estimate of PDF
+        evaluate           - compute value of PDF at given vector
+        plot_samples       - make triangle plot for observed data in M31 ref frame
+        set_PDF            - set the L.PDF field
+        test_gauss         - plots and calculates score vs ngauss components
+        preprocess_samples - zero the mean of the samples and scale by standard deviation
 
         NB. "vector" refes to a position in 6D MW-M31 D,vr,vt space
 
@@ -60,7 +62,10 @@ class Likelihood(object):
         self.T.transform_to_M31()
         #dt = np.dtype([('MW_D', 'f8'), ('MW_vr', 'f8'), ('MW_vt', 'f8'), ('M33_D', 'f8'), ('M33_vr', 'f8'), ('M33_vt', 'f8')])
         self.samples = np.transpose(np.array(self.T.get_kinematics()))
-        
+        self.samples_means = np.array([np.mean(self.samples[:,i]) for i in range(self.samples.shape[1])])
+        self.samples_stds = np.array([np.std(self.samples[:,i]) for i in range(self.samples.shape[1])])
+
+
 
 
         # PJM: Might be better to have Triplet.get_kinematics do this 
@@ -125,7 +130,13 @@ class Likelihood(object):
             bic_scores.append(pdf.bic(test_data))
         return aic_scores, bic_scores
 
+# ======================================================================
 
+    def preprocess_samples(self):
+
+        self.samples = (self.samples - self.samples_means)/self.samples_stds
+
+        return
 
 # ======================================================================
     def gaussianOverlay(self, figure):

### 2014-12-10 doc/ms1.tex
Placeholder for Marc's likelihood approximation section. Updated author list, it's by no means complete.

@@
 \author{Philip J. Marshall\altaffilmark{1}}
+\author{Marc Williamson\altaffilmark{1}}
 \author{Michael T. Busha\altaffilmark{2}} 
-\author{Risa H. Wechsler\altaffilmark{3, 4}}
+\author{Risa H. Wechsler\altaffilmark{1}}
 \author{Peter Behroozi\altaffilmark{3}}
-\author{XX\altaffilmark{5}}
-\author{YY\altaffilmark{6}}
-\altaffiltext{1}{Department of Physics, Oxford University, Oxford, OX1 3RH, UK.} 
-\altaffiltext{2}{Institute for Theoretical Physics, University of
- Zurich, 8057 Zurich, Switzerland}  
-\altaffiltext{3}{Kavli Institute for Particle Astrophysics and Cosmology,
-Physics Department, Stanford University, Stanford, CA, 94305, USA
-{\tt mbusha@physik.uzh.ch, pjm@slac.stanford.edu, rwechsler@stanford.edu}}
-\altaffiltext{4}{Particle and Particle Astrophysics Department, 
-SLAC National Accelerator Laboratory, Menlo Park, CA, 94025, USA}
-\altaffiltext{5}{XX}
-\altaffiltext{6}{YY}
+\author{XX\altaffilmark{4}}
+\author{YY\altaffilmark{5}}
+\altaffiltext{1}{Kavli Institute for Particle Astrophysics and Cosmology, Stanford, CA, 94305, USA
+{\tt pjm@slac.stanford.edu, rwechsler@stanford.edu}}
+\altaffiltext{2}{???}
+\altaffiltext{3}{STScI}
+\altaffiltext{4}{YY}
 
@@
 
-% ----------------------------------------------------------------------------
+% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
 
-\section{Observational Constraints}
+\subsection{Observational Constraints}
 \label{sec:data}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
+
+\subsection{Approximating the Likelihood}
+\label{sec:lhood}
+
+\todo{Marc}{Write this section!}
 

### 2014-12-23 doc/ms1.tex
Update ms1.tex

@@
 \author{Risa H. Wechsler\altaffilmark{1}}
-\author{Peter Behroozi\altaffilmark{3}}
-\author{XX\altaffilmark{4}}
-\author{YY\altaffilmark{5}}
+\author{XX\altaffilmark{3}}
+\author{YY\altaffilmark{4}}
 \altaffiltext{1}{Kavli Institute for Particle Astrophysics and Cosmology, Stanford, CA, 94305, USA
@@
 \altaffiltext{2}{???}
-\altaffiltext{3}{STScI}
+\altaffiltext{3}{XX}
 \altaffiltext{4}{YY}
@@
 
+We use the observed distances and radial and tangential velocities of the MW and M33 in Table 1 to construct a probability density function for the Local Group.
+Assuming that the values in the table come from Gaussian distributions, we sample a large number of systems similar to the Local Group.
+
+The density of this "observational" dataset is fit well by a Gaussian mixture model (GMM).
+(Plot showing goodness of fit?
+Triangular plot?) The optimal number of components in the GMM is found using the Bayesian Information Criterion with cross validation.
+The GMM allows us to evaluate the likelihood of a Consuelo Local Group under the observational pdf.
+This likelihood is an importance weight which allows us to find marginalized posterior pdf's for the Consuelo halo properties.
+
 % ----------------------------------------------------------------------------

### 2014-12-24 localgroup/triplet.py
debugging deltavrot

@@ -148,7 +148,14 @@ if __name__ == '__main__':
     print np.median(D_M33), np.median(vr_M33), np.median(vt_M33)
 
 
+
+
+
+
+
+
     """
+    
     print "Checking MW:"
     # Check speed:
     w = np.sqrt(t.MW.vx[0]*t.MW.vx[0] + t.MW.vy[0]*t.MW.vy[0] + t.MW.vz[0]*t.MW.vz[0])
@@ -171,6 +178,8 @@ if __name__ == '__main__':
     print "M33 speeds: ",u,w
 
     print " "
+    """
+
     
     print "Before taking out solar motion, M31 velocity: ",np.mean(t.M31.vx),'+/-',np.std(t.M31.vx),', ', \
                            np.mean(t.M31.vy),'+/-',np.std(t.M31.vy),', ', \
@@ -190,10 +199,10 @@ if __name__ == '__main__':
     print "M31 position: ",np.mean(t.M31.x),np.mean(t.M31.y),np.mean(t.M31.z)
     print "  cf vdM++12: (-0.379, 0.613, -0.283)"
     
-    print "M31 proper motion: ",np.mean(t.M31.v_west),'+/-',np.std(t.M31.v_west),', ', \
-                                np.mean(t.M31.v_north),'+/-',np.std(t.M31.v_north)
-    print "  cf vdM++12: (-125+/-31, -74+/-28) km/s"
-    print "This is a little off because of our deltavrot hackery..."
+ #   print "M31 proper motion: ",np.mean(t.M31.v_west),'+/-',np.std(t.M31.v_west),', ', \
+  #                              np.mean(t.M31.v_north),'+/-',np.std(t.M31.v_north)
+   # print "  cf vdM++12: (-125+/-31, -74+/-28) km/s"
+   # print "This is a little off because of our deltavrot hackery..."
    
     print " "
     
@@ -207,4 +216,4 @@ if __name__ == '__main__':
     print "  cf vdM++12: (110.6 +/- 7.8) km/s"
 
     print "Looks like there's a bug in the Galactocentric velocity transformation..."
-    """
+  

### 2015-03-15 localgroup/triplet.py
saves sim Mvir values

@@ -131,6 +131,7 @@ class Triplet(object):
         self.observe_halos(Nsamples=n_points)
 
         self.MW.translate_to(self.MW)
+        self.MW.Mvir = self.sim_data['MW_Mvir']
 
         self.M31.x = self.sim_data['M31_x'] - self.sim_data['MW_x']
         self.M31.y = self.sim_data['M31_y'] - self.sim_data['MW_y']
@@ -139,6 +140,7 @@ class Triplet(object):
         self.M31.vy = self.sim_data['M31_vy'] - self.sim_data['MW_vy']
         self.M31.vz = self.sim_data['M31_vz'] - self.sim_data['MW_vz']
         self.M31.frame = 'MW'        
+        self.M31.Mvir = self.sim_data['M31_Mvir']
         if not self.isPair:
             self.M33.x = self.sim_data['M33_x'] - self.sim_data['MW_x']
             self.M33.y = self.sim_data['M33_y'] - self.sim_data['MW_y']
@@ -147,7 +149,7 @@ class Triplet(object):
             self.M33.vy = self.sim_data['M33_vy'] - self.sim_data['MW_vy']
             self.M33.vz = self.sim_data['M33_vz'] - self.sim_data['MW_vz']
             self.M33.frame = 'MW'
-
+            self.M33.Mvir = self.sim_data['M33_Mvir']
         return
 # ============================================================================
 
@@ -162,6 +164,13 @@ class Triplet(object):
         total_weight = weights.sum()
         self.weights = 1.0/total_weight*weights
         return
+
+# ============================================================================
+
+
+
+
+
 # ============================================================================
     def calculate_N95(self):
         weights_copy = np.copy(self.weights)

### 2015-03-17 main.tex

\documentclass[iop,apj]{emulateapj}
\usepackage{amsmath,amssymb,amstext}

\usepackage[breaklinks,colorlinks,citecolor=blue,linkcolor=magenta]{hyperref} 
\renewcommand*{\sectionautorefname}{Section}
\usepackage[all]{hypcap} %Links go to figures; breaks on deluxetables (use \capstartfalse \capstarttrue to fix it)

\usepackage{aas_macros}
\usepackage{natbib}
\bibliographystyle{apj}

\shorttitle{Short Title}
\shortauthors{Author et al.}

\begin{document}

\title{Template for the Astrophysical Journal with emulateapj}
\author{Authors}
\affil{Affiliations}

\begin{abstract}
Abstract.
\end{abstract}

\keywords{keywords}
\maketitle

\acknowledgments{
  Acknowledgments. 
}

\end{document}

### 2015-03-30 localgroup/__init__.py
before merging with Phil's label work

@@ -4,3 +4,4 @@ from triplet import *
 from sampling import *
 from coordinates import *
 from likelihood import *
+#from coords import *

### 2015-03-30 localgroup/coordinates.py

@@ -65,7 +65,7 @@ def equatorial_to_galactic_proper_motion(mu_w, mu_n,RA,DEC):
 # mu_[west|north] in mas/yr, v_r in km/s
 # x,y,z in Mpc.  vx, vy, vz, in km/s
 
-def spherical_to_cartesian(RA,DEC,D,mu_west,mu_north,v_r,deltavrot_west, deltavrot_north):
+def spherical_to_cartesian(RA,DEC,D,mu_west,mu_north,v_r,deltavrot_west, deltavrot_north, M31):
 
     # Heliocentric Cartesian positions (Mpc):
     delta = DEC*deg2rad
@@ -75,8 +75,12 @@ def spherical_to_cartesian(RA,DEC,D,mu_west,mu_north,v_r,deltavrot_west, deltavr
     z = D*sin(delta)
      
     # Convert mu to v
-    v_west = -mu_west*muaspyrMpc2kmps*D + 1*deltavrot_west
-    v_north = mu_north*muaspyrMpc2kmps*D - 1*deltavrot_north
+    if not M31:
+        v_west = -mu_west*muaspyrMpc2kmps*D + 1*deltavrot_west
+        v_north = mu_north*muaspyrMpc2kmps*D - 1*deltavrot_north
+    else:
+        v_west = -mu_west
+        v_north = mu_north
     print 'v_west:  %f +/- %f'%(mean(v_west),std(v_west)) 
     print 'v_north: %f +/- %f'%(mean(v_north),std(v_north))
  
@@ -105,7 +109,7 @@ def heliocentric_galactic_cartesian_to_galactocentric_cartesian(xh,yh,zh,vxh,vyh
 # galactocentric cartesian coordinates (x,y,z, etc).  
 # NOTE:  Add optional argument to take care of internal rotation.  
 
-def heliocentric_equatorial_spherical_to_galactocentric_cartesian(ra, dec, d, mu_w, mu_n, v_r, dvrot_w, dvrot_n, R0=0.00829, VX=-11, V0=-239, VZ=-7):
+def heliocentric_equatorial_spherical_to_galactocentric_cartesian(ra, dec, d, mu_w, mu_n, v_r, dvrot_w, dvrot_n, R0=0.00829, VX=-11, V0=-239, VZ=-7, M31=True):
[…]

### 2015-03-30 localgroup/halo.py

@@ -102,7 +102,9 @@ class Halo(object):
             self.mu_north        = localgroup.draw(obs['mu_north'],Nsamples)
             self.deltavrot_north = localgroup.draw(obs['deltavrot_north'],Nsamples)
             self.v_r             = localgroup.draw(obs['v_r'],Nsamples)
-
+            if self.name == 'M31':
+                self.v_west = localgroup.draw(obs['v_west'], Nsamples)
+                self.v_north = localgroup.draw(obs['v_north'], Nsamples)
 
             # BUG!  Transformation here may be redundant with functions defined in coordinates.py!  Eliminate redundancies!
             # Equatorial tangential motions in km/s:

### 2015-03-30 localgroup/likelihood.py

@@ -170,7 +170,7 @@ class Likelihood(object):
         else:
             size = 6
         n_gaussians = ngauss
-        if ngauss > 5: raise AttributeError("Only 5 colors can be shown.")
+        #if ngauss > 5: raise AttributeError("Only 5 colors can be shown.")
         colors = ['g', 'r', 'y', 'b', 'c']
         transparency = 0.5
         model = mixture.GMM(n_gaussians, covariance_type='full')
@@ -185,7 +185,7 @@ class Likelihood(object):
                         mean = [model.means_[gauss_num][j], model.means_[gauss_num][i]]
                         covar = model.covars_[gauss_num]
                         covar = [[covar[j,j], covar[j,i]], [covar[i,j], covar[i,i]]]
-                        color = colors[gauss_num]
+                        color = colors[gauss_num%5]
                         v, w = linalg.eigh(covar)
                         u = w[0]/linalg.norm(w[0])
                         angle = np.arctan(u[1]/u[0])

### 2015-03-30 localgroup/triplet.py

@@ -4,9 +4,12 @@ import localgroup
 import triangle
 import numpy as np
 import sys
-sys.path.append('/afs/slac.stanford.edu/u/ki/yymao/scripts')
-from readHlist import readHlist
+#sys.path.append('/afs/slac.stanford.edu/u/ki/yymao/scripts')
+#from readHlist import readHlist
 import pickle
+
+sys.path.append('/u/ki/yymao/pyscripts')
+from helpers.SimulationAnalysis import readHlist
 # ======================================================================
 
 class Triplet(object):
@@ -101,10 +104,10 @@ class Triplet(object):
         # Covert M31 and M33 from heliocentric spherical to 
         # galactocentric cartesian.
         if not sim:  
-            self.M31.x, self.M31.y, self.M31.z, self.M31.vx, self.M31.vy, self.M31.vz = localgroup.heliocentric_equatorial_spherical_to_galactocentric_cartesian(self.M31.RA,self.M31.DEC,self.M31.D,self.M31.mu_west,self.M31.mu_north,self.M31.v_r, self.M31.deltavrot_west, self.M31.deltavrot_north, R0=self.MW.x, VX=self.MW.vx, V0=self.MW.vy, VZ=self.MW.vz)
+            self.M31.x, self.M31.y, self.M31.z, self.M31.vx, self.M31.vy, self.M31.vz = localgroup.heliocentric_equatorial_spherical_to_galactocentric_cartesian(self.M31.RA,self.M31.DEC,self.M31.D,self.M31.v_west,self.M31.v_north,self.M31.v_r, self.M31.deltavrot_west, self.M31.deltavrot_north, R0=self.MW.x, VX=self.MW.vx, V0=self.MW.vy, VZ=self.MW.vz, M31=True)
             self.M31.frame = 'MW'
             if not self.isPair:
-                self.M33.x, self.M33.y, self.M33.z, self.M33.vx, self.M33.vy, self.M33.vz = localgroup.heliocentric_equatorial_spherical_to_galactocentric_cartesian(self.M33.RA,self.M33.DEC,self.M33.D,self.M33.mu_west,self.M33.mu_north,self.M33.v_r, self.M33.deltavrot_west, self.M33.deltavrot_north, R0=self.MW.x, VX=self.MW.vx, V0=self.MW.vy, VZ=self.MW.vz)
[…]

### 2015-03-30 scripts/compile_all_M33_data.py

@@ -1,15 +1,16 @@
 import numpy as np
 import sys
-sys.path.append('/afs/slac.stanford.edu/u/ki/yymao/scripts')
-from readHlist import readHlist
+
+sys.path.append('/u/ki/yymao/pyscripts')
+from helpers.SimulationAnalysis import readHlist
 
 header = '# idMW idM31 idM33 Mvir x y z vx vy vz\n'
 
 path = '/nfs/slac/g/ki/ki04/mbusha/projects/LasDamas/Simulations/Consuelo/LG_project/'
-M33_prefix = 'M33_halos_'
+M33_prefix = 'MW_halos_'
 M33_suffix = '_z0p000_vir_bgc2.txt'
 
-M33_file = open('M33_all_z0p000_data.txt','a+')
+M33_file = open('MW_all_z0p000_data.txt','a+')
 M33_file.write(header)
 for i in range(4001,4051):
     M33_file_name = path+M33_prefix+str(i)+M33_suffix

### 2015-04-04 localgroup/coords.py
updating organization

@@ -1,0 +1,196 @@
+# ======================================================================
+
+from numpy import mean,sin,cos,arcsin,arccos,arctan2,pi,abs,sqrt,sign,atleast_1d,arctan,std
+from pytpm import tpm, convert
+import numpy as np
+import math
+deg2rad = pi/180.0
+arcsec2rad = deg2rad/3600.0
+muaspyrMpc2kmps = 4.7404
+
+# North Galactic pole:
+alphaG = 192.85948 * deg2rad
+deltaG =  27.12825 * deg2rad
+# Galactic longitude of the ascending node of the Galactic plane:
+lOmega = 32.93192 * deg2rad
+
+# Numerical tolerance:
+tol = 1e-8
+
+# ======================================================================
+# Rotate sky positions from equatorial to galactic coordinates.
+# From Poleski 2013.
+
+def equatorial_to_galactic(RA,DEC):
+    
+    alpha = RA*deg2rad
+    delta = DEC*deg2rad
+    
+    b = arcsin(cos(delta)*cos(deltaG)*cos(alpha - alphaG) + sin(delta)*sin(deltaG))
+    b = atleast_1d(b)
+
+    sindl = cos(delta)*sin(alpha-alphaG)/cos(b)
+    cosdl = (sin(delta)*cos(deltaG) - cos(delta)*sin(deltaG)*cos(alpha-alphaG))/cos(b)
+    dl = arctan2(cosdl,sindl)
+    l = atleast_1d(lOmega + dl)
+    l[l < 0.0] += 2.0*pi
+   
+ 
+    return l/deg2rad,b/deg2rad
+    
+# ----------------------------------------------------------------------
+# Generic transformation between spherical and cartesian coordinates.
+# For galactic coords, use l,b instead of RA,DEC, and so on.
+# UNITS:  RA, DEC should be in degrees.  D is in Mpc.  
+# mu_[west|north] in mas/yr, v_r in km/s
+# x,y,z in Mpc.  vx, vy, vz, in km/s
+
+def spherical_to_cartesian(RA,DEC,D,mu_west,mu_north,v_r,deltavrot_west, deltavrot_north, M31):
+
+    # Heliocentric Cartesian positions (Mpc):
+    delta = DEC*deg2rad
+    alpha = RA*deg2rad
+    x = D*cos(delta)*cos(alpha)
+    y = D*cos(delta)*sin(alpha)
[…]

### 2015-04-04 localgroup/likelihood.py

@@ -149,7 +149,6 @@ class Likelihood(object):
 
         if overlay:
             self.gaussianOverlay(figure, ngauss)
-        figure.savefig("L_samples_tri.png")
 
         return figure
 

### 2015-04-04 doc/thesis/main.tex

--- 2014-12-23 doc/ms1.tex
@@
-\documentclass{emulateapj}
+\documentclass[iop,apj]{emulateapj}
+\usepackage{amsmath,amssymb,amstext}
 
-\usepackage{apjfonts}
-\usepackage{epsfig}
+%\usepackage[breaklinks,colorlinks,citecolor=blue,linkcolor=magenta]{hyperref} 
+%\renewcommand*{\sectionautorefname}{Section}
+%\usepackage[all]{hypcap} %Links go to figures; breaks on deluxetables (use \capstartfalse \capstarttrue to fix it)
+
+\usepackage{aas_macros}
 \usepackage{natbib}
+\bibliographystyle{apj}
+
+\shorttitle{Short Title}
+\shortauthors{Author et al.}
 
@@
 \citestyle{aa}
-\shortauthors{Marshall et al.}
+\shortauthors{Williamson et al.}
 \shorttitle{The Mass of the Local Group}
@@
 
+\author{Marc Williamson\altaffilmark{1}}
 \author{Philip J. Marshall\altaffilmark{1}}
-\author{Marc Williamson\altaffilmark{1}}
 \author{Michael T. Busha\altaffilmark{2}} 
 \author{Risa H. Wechsler\altaffilmark{1}}
-\author{XX\altaffilmark{3}}
-\author{YY\altaffilmark{4}}
+
 \altaffiltext{1}{Kavli Institute for Particle Astrophysics and Cosmology, Stanford, CA, 94305, USA
-{\tt pjm@slac.stanford.edu, rwechsler@stanford.edu}}
+{\tt mwillia1@stanford.edu, pjm@slac.stanford.edu, rwechsler@stanford.edu}}
 \altaffiltext{2}{???}
-\altaffiltext{3}{XX}
-\altaffiltext{4}{YY}
 
@@
 at the redshift under consideration.
+
+%-----------------------------------------------------------------------------
+
+\section{Coordinates}
+\label{sec:coord}
+
+In this section we describe the coordinate systems used in this paper and present the transformations between them.
+As mentioned in the introduction, this project uses the M31 galactocentric dynamics of the MW-M31-M33 system to define a LG analog.
+Specifically, we use the parameters $\distance$, $\vrad$, and $\vtan$ measured from M31 to MW and M33 respectively.
+The LG analog sample is drawn using the "observed" values of these parameters.
+However, the true observed values of these parameters are of course geocentric and measured in equatorial coordinates of right ascension ($\alpha$) and declination ($\delta$).
+Therefore, it is necessary to be able to transform from this system to the M31 galactocentric coordinate system.
+This involves many intermediate coordinate systems and transformations which will be described in the remainder of this section.
+\subsection{Coordinate System Definitions}
+\label{sec:c_sys_defs}
+Three types of coordinate systems are used in this work: Geocentric Equatorial, Heliocentric Galactic, and Galactocentric.
+Before we describe the mathematical transformations between these systems, it is important to thoroughly understand the definitions of the systems.
+\par The equatorial system is centered on the Earth and based on the concept of a celestial sphere (ie all stars are projected onto a sphere of infinite radius centered on the Earth).
+Two angles characterize the positions of objects on the celestial sphere: right ascension and declination.
+Right ascension, $\alpha$, is the angle measured eastwards along the celestial equator between the vernal equinox and the hour circle passing through an object.
+Declination, $\delta$, is the angular distance of an object above or below the celestial equator.
+\par The heliocentric galactic coordinate system (often ambiguously referred to as galactic coordinates), is a different spherical coordinate system centered at the Sun.
+The heliocentric fundamental plane is aligned with the galactic plane, and the primary direction is along the line connecting the sun and the galactic center.
+Positions are measured in galactic latitude and longitude.
+Latitude, $b$, measures angular distance of an object perpendicular to the galactic plane, with positive angles to the north galactic pole.
+Longitude, $l$, is the angular distance measured eastward along the galactic equator from the galactic center.
+\par The galactocentric coordinate system used in this paper is the same as the one used by van der Marel \citep{VdM08}.
+The $Z$ axis points towards the galactic north pole, the $Y$ axis is oriented in the direction of the sun's rotation about the galactic center, and the $X$ axis points from the sun towards the galactic center.
+This is a standard, right-handed coordinate system.
 
@@
 % ----------------------------------------------------------------------------
+%\nocite{*}
+%\bibliographystyle{unsrt}
+%\bibliography{references.bib}
 
-\bibliographystyle{apj}
+% ----------------------------------------------------------------------------
+\nocite{*}
 \bibliography{references}
 
-% ----------------------------------------------------------------------------
-
 \end{document}

### 2015-04-06 localgroup/likelihood.py
triangle overlay uses pappy

@@ -64,10 +64,10 @@ class Likelihood(object):
 
         if self.T.isPair:
             # self.labs = ["MW_D", "MW_vr", "MW_vt"]
-            self.labels = ["$\Delta D^{\\rm MW}$", "$\Delta v_{\\rm rad}^{\\rm MW}$", "$\Delta v_{\\rm tan}^{\\rm MW}$"]
+            self.labels = ["$D^{\\rm MW}$", "$v_{\\rm rad}^{\\rm MW}$", "$v_{\\rm tan}^{\\rm MW}$"]
         else:
             # self.labs = ["MW_D", "MW_vr", "MW_vt", "M33_D", "M33_vr", "M33_vt"]
-            self.labels = ["$\Delta D^{\\rm MW}$", "$\Delta v_{\\rm rad}^{\\rm MW}$", "$\Delta v_{\\rm tan}^{\\rm MW}$", "$\Delta D^{\\rm M33}$", "$\Delta v_{\\rm rad}^{\\rm M33}$", "$\Delta v_{\\rm tan}^{\\rm M33}$"]
+            self.labels = ["$D^{\\rm MW}$", "$v_{\\rm rad}^{\\rm MW}$", "$v_{\\rm tan}^{\\rm MW}$", "$D^{\\rm M33}$", "$v_{\\rm rad}^{\\rm M33}$", "$v_{\\rm tan}^{\\rm M33}$"]
 
         return
 
@@ -122,26 +122,25 @@ class Likelihood(object):
 
 # ======================================================================
 
-    def model_gof(self, n_points, mode="GMM"):
-
+    def model_gof(self, n_points, color, fig=None, mode="GMM"):
         self.write_labels()
 
         if (mode == "GMM"):
-            drawn_points = self.PDF.sample(n_samples=n_points)
-            figure = triangle.corner(drawn_points, labels=self.labels, quantiles=[0.16,0.5,0.84], show_titles=True, title_args={"fontsize": 12})
+            drawn_points = self.PDF.sample(n_samples=n_points)*self.samples_stds + self.samples_means
+            figure = triangle.corner(drawn_points, labels=self.labels, quantiles=[0.16,0.5,0.84], fig=fig, show_titles=True, title_args={"fontsize": 12}, color=color)
         else:
             raise ValueError("Unrecognized approximation mode %s" % mode)
-
-        return figure
+        if fig==None: return figure
+        return fig
 
 # ======================================================================
 
[…]

### 2015-04-06 localgroup/triplet.py

@@ -185,22 +185,29 @@ class Triplet(object):
         self.sim_samples = (self.sim_samples - means)/stds
 
         return
+
 # ============================================================================
 
-    def plot_kinematics(self):
+    def unprocess(self, means, stds):
+        self.sim_samples = self.sim_samples*stds + means
+        return
 
+# ============================================================================
+
+    def plot_kinematics(self, means, stds, color, fig=None):
+        self.unprocess(means, stds)
         if self.isPair:
             # labs = ["MW_D", "MW_vr", "MW_vt"]
-            labels = ["$\Delta D^{\\rm MW}$", "$\Delta v_{\\rm rad}^{\\rm MW}$", "$\Delta v_{\\rm tan}^{\\rm MW}$"]
+            labels = ["$D^{\\rm MW}$", "$v_{\\rm rad}^{\\rm MW}$", "$v_{\\rm tan}^{\\rm MW}$"]
         else:
             # labs = ["MW_D", "MW_vr", "MW_vt", "M33_D", "M33_vr", "M33_vt"]
-            labels = ["$\Delta D^{\\rm MW}$", "$\Delta v_{\\rm rad}^{\\rm MW}$", "$\Delta v_{\\rm tan}^{\\rm MW}$", "$\Delta D^{\\rm M33}$", "$\Delta v_{\\rm rad}^{\\rm M33}$", "$\Delta v_{\\rm tan}^{\\rm M33}$"]
+            labels = ["$D^{\\rm MW}$", "$v_{\\rm rad}^{\\rm MW}$", "$v_{\\rm tan}^{\\rm MW}$", "$D^{\\rm M33}$", "$v_{\\rm rad}^{\\rm M33}$", "$v_{\\rm tan}^{\\rm M33}$"]
 
         if self.isPair:
-            figure = triangle.corner(self.sim_samples, labels=labels, quantiles=[0.16,0.5,0.84], show_titles=True, title_args={"fontsize": 12})
+            figure = triangle.corner(self.sim_samples, labels=labels, quantiles=[0.16,0.5,0.84], fig=fig, show_titles=True, title_args={"fontsize": 12}, color=color)
         else:
-            figure = triangle.corner(self.sim_samples, labels=labels, quantiles=[0.16,0.5,0.84], show_titles=True, title_args={"fontsize": 12})
-
+            figure = triangle.corner(self.sim_samples, labels=labels, quantiles=[0.16,0.5,0.84], fig=fig, show_titles=True, title_args={"fontsize": 12}, color=color)
[…]

### 2015-04-06 scripts/compile_all_M33_data.py
made pdf plots and posterior plots

@@ -7,10 +7,10 @@ from helpers.SimulationAnalysis import readHlist
 header = '# idMW idM31 idM33 Mvir x y z vx vy vz\n'
 
 path = '/nfs/slac/g/ki/ki04/mbusha/projects/LasDamas/Simulations/Consuelo/LG_project/'
-M33_prefix = 'MW_halos_'
+M33_prefix = 'M31_halos_'
 M33_suffix = '_z0p000_vir_bgc2.txt'
 
-M33_file = open('MW_all_z0p000_data.txt','a+')
+M33_file = open('M31_all_z0p000_data.txt','a+')
 M33_file.write(header)
 for i in range(4001,4051):
     M33_file_name = path+M33_prefix+str(i)+M33_suffix

### 2015-04-06 scripts/compile_all_complete_triplets.py

@@ -3,21 +3,24 @@
 
 import numpy as np
 import sys
-sys.path.append('/afs/slac.stanford.edu/u/ki/yymao/scripts')
-from readHlist import readHlist
+import os
+sys.path.append('/u/ki/yymao/pyscripts')
+from helpers.SimulationAnalysis import readHlist
 
 # idMW idM31 idM33 Mvir x y z vx vy vz
 header = '# idMW idM31 idM33 MW_Mvir M31_Mvir M33_Mvir MW_x MW_y MW_z M31_x M31_y M31_z M33_x M33_y M33_z\
  MW_vx MW_vy MW_vz M31_vx M31_vy M31_vz M33_vx M33_vy M33_vz\n'
-path = '/afs/slac.stanford.edu/u/ki/mwillia1/Thesis/data_files/'
+path = '/afs/slac.stanford.edu/u/ki/mwillia1/Thesis/LocalGroupHaloProps/scripts/'
 M33_file = 'M33_all_z0p000_data.txt'
-MW_trip_file = 'MW_all_Triplets_z0p000_data.txt'
-M31_trip_file = 'M31_all_Triplets_z0p000_data.txt'
+MW_trip_file = 'MW_all_z0p000_data.txt'
+M31_trip_file = 'M31_all_z0p000_data.txt'
 
-M33_data = readHlist(path+M33_file)
+#M33_data = readHlist(path+M33_file)
 MW_data = readHlist(path+MW_trip_file)
+os.system('rm %s'%(path+MW_trip_file))
+os.system('python compile_all_M33_data.py')
 M31_data = readHlist(path+M31_trip_file)
-
+os.system('rm %s'%(path+M31_trip_file))
 
 triplet_file = open('complete_triplets.txt', 'w')
 triplet_file.write(header)
@@ -25,8 +28,10 @@ for MW_line in MW_data:
     MW_id = MW_line['idMW']
     M31_id = MW_line['idM31']
     M33_id = MW_line['idM33']
+    if not M33_id==-1: continue
     M31_line = M31_data[M31_data['idM31']==M31_id][0]
-    M33_line = M33_data[M33_data['idM33']==M33_id][0]
+    #M33_line = M33_data[M33_data['idM33']==M33_id][0]
+    M33_line = np.copy(M31_line)
     triplet_file.write('%d %d %d %f %f %f '%(MW_line['idMW'], M31_line['idM31'], M33_line['idM33'], \
[…]

### 2015-04-10 doc/thesis/main.tex
Updated abstract, intro

@@
 \usepackage{natbib}
+%\usepackage{siunitx}
 \bibliographystyle{apj}
@@
 
-We study the local group of galaxies in its cosmological context, by matching
-dark matter halos in the \consuelo N-body simulation with kinematic
-observations of M31 and M33. Treating the \consuelo halo  catalog as a list of
-samples drawn from the cosmological prior PDF for the local group properties,
-we select local group analogs whose isolation matches our own, and then
-compute importances for the samples as combined likelihoods for the
-observations. 
-%
-From the ``observed'' galactocentric distance, radial velocity and tangential 
-velocity of M31 alone, we infer the mass of the local group to be $\MLG =
+We study the local group of galaxies by identifying analogous systems in the \consuelo N-body simulation.
+The \consuelo halo catalogs are treated as a large sample drawn from the cosmological prior PDF, and local group analogs are identified by matching the observed kinematics of the system.
+Weights are calculated for the \consuelo sample by finding the likelihoods for the simulated kinematic properties compared to observations.
+We study the effect of including the existence and kinematics of M33 on the posterior distributions of the local group mass by using three types of samples: from only the "observed" galactocentric distance, radial velocity, and tangential velocity of M31, we infer the mass of the local group to be $\MLG =
+(\MMW + \MEI) = (\MPAIRestimate^{+\MPAIRerrorplus}_{-\MPAIRerrorminus}) \times
+10^{12} \Msun$. Including the only the existence of M33 as a condition for identifying local group analogs results in a different inference of the mass, $\MLG =
+(\MMW + \MEI) = (\MPAIRestimate^{+\MPAIRerrorplus}_{-\MPAIRerrorminus}) \times
+10^{12} \Msun$. Finally including the kinematics of M33 we infer the mass of the local group to be $\MLG =
 (\MMW + \MEI) = (\MPAIRestimate^{+\MPAIRerrorplus}_{-\MPAIRerrorminus}) \times
@@
 timing argument: we show that for our matched analogs the timing argument mass
-systematically ... the local group mass. 
-%
-Transforming the ``observed'' galactocentric distance, radial velocity and
-tangential  velocity of M33 into the M31 rest-frame, and accounting for
-covariance between these measurements, we also infer the mass of the M31--M33
-system to be  $(\MEI + \MEE) =
-(\MbAestimate^{+\MbAerrorplus}_{-\MbAerrorminus}) \times 10^{12} \Msun$.
-%
-Combining the kinematic constraints from both M31 and M33,  we infer a local
-group mass of $\MLG = (\MMW + \MEI) =
-(\MTRIPLETestimate^{+\MTRIPLETerrorplus}_{-\MTRIPLETerrorminus}) \times
-10^{12} \Msun$, and are able to estimate the halo masses of the three galaxies
+systematically ... the local group mass.  We are also able to estimate the halo masses of the three galaxies
 individually, finding these to be $\MEI =
@@
 
-Our goal is to map out the dark matter distribution in the Local Group
+Our goal is to tightly constrain the mass of the Local Group
 by combining observational data on the  positions and velocities of the
@@
 samples drawn from a very informative, highly correlated prior PDF for
-halo properties, and then  importance sample them with likelihood
+halo properties, and then  calculate weighted posterior PDF's using likelihood
 functions based on the observed kinematics of the particular system
@@
 \item How accurate is the Timing Argument for the Local Group mass?
+\item How do the existence and kinematics of M33 impact the Local Group mass?
 \item What are the masses of the dark matter halos of M31, M33 and the Milky
 Way?
+\item How sensitive is this method to the isolation criteria used to define an LG-analog?
 \end{itemize}
 
-This paper is structured as follows. In \Sref{sec:method} we review the
+This paper is structured as follows. In \Sref{sec:coord} we review the coordinate systems and transformations used in this project.  In \Sref{sec:method} we review the
 halo inference methodology, and introduce the \consuelo N-body
@@
 This is a standard, right-handed coordinate system.
+\subsection{Equatorial to Galactic Transformation}
+\label{sec:eq_to_gal}
+\par This transformation is a rotation from equatorial coordinates to galactic coordinates.
+Given the right ascension and declination of a galaxy, $(\alpha,\delta)$, the transformation returns the galactic latitude and longitude, $(l,b)$.
+We use the same method described by \cite{poleski2013}, solving the following system of equations to find the galactic coordinates:
+\begin{equation}
+\begin{aligned}
+\sin b &= \cos\delta\cos\delta_{G}\cos(\alpha - \alpha_{G}) + \sin\delta\sin\delta_{G} \\
+\sin(l_{\Omega} - l)\cos b &= \cos\delta\sin(\alpha - \alpha_{G}) \\
+\cos(l_{\Omega} - l)\cos b &= \sin\delta\cos\delta_{G} - \cos\delta\sin\delta_{G}\cos(\alpha - \alpha_{G})
+\end{aligned}
+\end{equation}
 
+where $\alpha_{G} = 192.85948$ and $\delta_{G} = 27.12825$ are the equatorial coordinates of the galactic north pole, and $l_{\Omega}$ is the galactic longitude of the ascending node of the galactic plane.
+\subsection{Galactic Spherical to Cartesian}
+\label{sec:sphere_to_cart}
+Before transforming to the galactocentric reference frame, we first transform the galactic latitude and longitude to cartesian coordinates.
+Expressing the coordinates in radians, the position in cartesians is simple:
+\begin{equation}
+\begin{aligned}
+x &= D\cos b\sin l \\
+y &= D\cos b\sin l \\
+z &= D\sin b
+\end{aligned}
+\end{equation}
+where $D$ is the distance to the observed galaxy (M31 or M33). 
+\par The velocity transformation is more complicated.  They are given by:
+\begin{equation}
+\begin{aligned}
+v_{x} &= v_{r}\cos(b)cos(l) - v_{W}\sin(l) - v_{N}\sin(b)\cos(l)\\
+v_{y} &= v_{r}\cos(b)\sin(l) + v_{W}\cos(l) - v_{N}\sin(b)\sin(l) \\
+v_{z} &= v_{r}\sin(b) + v_{N}\cos(b)
+\end{aligned}
+\end{equation}
+where $v_{r}$ is the line of sight velocity of the observed galaxy, and $(v_{W},v_{N})$ are the linear velocities in the west and north directions as measured in the equatorial frame.
+In our project, we treat the velocity transformations of M31 and M33 differently due to better constraints on $(v_{W},v_{N})$ for M31 existing in the literature.
+For M31, we use van der Marel's constraints as the "observed" values: $(v_{W},v_{N}) = (-125.2 \pm 30.8, -73.8 \pm 28.4)$ \citep{vdm12m31}.
+There are only observations of the proper motion of M33 in spherical equatorial coordinates, so an extra transformation must be made to find the $(v_{W},v_{N})$ for M33.
+We follow van der Marel's method exactly from \cite{VdM08}.
+\subsection{Heliocentric to Galactocentric}
+\label{sec:hel_to_gal}
+Transforming to the Galactocentric cartesian frame from the Heliocentric cartesian frame is a simple translation in the $X$ direction. The galactocentric positions are thus:
+\begin{equation}
+\begin{aligned}
+x_{g} &= x_{h} - R_{0}
+y_{g} &= y_{h} \\
+z_{g} &= z_{h}
+\end{aligned}
+\end{equation}
+where $R_{0} = 8.29 \pm 0.16$ kpc as used in \cite{vdm12m31}.
+The velocity transformations are slightly more complicated than a normal translation because the effects of the solar velocity in the central galaxy must be removed.
+We use the values derived by Schönrich et al: $v_{\odot} = (11.1, 12.24, 7.25)$ with the listed uncertainties \citep{sunlsr}.
+Thus the velocities in the galactocentric frame are:
+\begin{equation}
+\begin{aligned}
+v_{x,g} &= v_{x,h} - v_{\odot,x} \\
+v_{y,g} &= v_{y,h} - (V_{0}+v_{\odot,y})\\
+v_{z,g} &= v_{z,h} - v_{\odot,z}
+\end{aligned}
+\end{equation}
+where $V_{0} = 239.3 \pm 10.3$ is the circular velocity of the local standard of rest about the galactic center \citep{vdm12m31}.
 % ----------------------------------------------------------------------------
@@
 
-\todo{Phil}{Write this section!}
+\todo{Marc}{Write this section!}
 
@@
 \subsection{The \consuelo Simulation}
-\label{sec:sampling}
+\label{sec:sampling_b}
 
-\todo{Mike}{Write this section!}
+\todo{Marc}{Write this section!}
 
@@
 \subsection{The Local Group prior}
-\label{sec:sampling}
+\label{sec:sampling_c}
 
-\todo{Mike}{Write this section!}
+\todo{Marc}{Write this section!}
 
@@
 
+\begin{figure*}
+  \centering \includegraphics[width=\textwidth]{plots/model_gof.png}
+  \caption{The 1-D and 2-D marginalizations of the parameter space used to create the GMM model.
+The sample generated from the observed values of the 6 parameters (blue) and the sample drawn from the resulting GMM (red) overlap almost perfectly.}
+  \label{fig:gof}
+\end{figure*}
+
+\begin{figure*}
+  \centering \includegraphics[width=\textwidth]{plots/sim_plot.png}
+  \caption{The 1-D and 2-D marginalizations of the CONSUELO triplet sample (green) overlayed with the corresponding marginalizations from the sample drawn from the GMM model (red).
+The intersection shows which of the CONSUELO triplets are given the most weight by the model.}
+  \label{fig:sim_plot}
+\end{figure*}
+
 % ----------------------------------------------------------------------------
@@
 Corner plot of Pr(MMW, MM31, MLG), prior vs all constraints
+
+\begin{figure}
+\includegraphics[width=\linewidth]{plots/res_plot.png}
+  \caption{Log masses (units of $M_{\odot}$) of the MW, M31, and the LG weighted by likelihood. The displayed values above the histograms are the triplet numbers.}
+  \label{fig:res_plot}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\linewidth]{plots/pdf_lg_plot.png}
+  \caption{The likelihood distribution of the mass of the Local Group. The Triplet distribution suffers from small statistics.}
+  \label{fig:lg_pdf}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\linewidth]{plots/pdf_mw_plot.png}
+  \caption{The likelihood distribution of the mass of the MW. The Triplet distribution suffers from small statistics.}
+  \label{fig:mw_pdf}
+\end{figure}
 

### 2015-08-24 doc/paper/main.tex
Checked in files from overleaf repo

\documentclass[iop,apj,twocolappendix,numberedappendix]{emulateapj}
\usepackage{amsmath,amssymb,amstext}

\usepackage[breaklinks,colorlinks,citecolor=blue,linkcolor=magenta]{hyperref}
\usepackage[all]{hypcap} %Links go to figures; breaks on deluxetables (use \capstartfalse \capstarttrue to fix it)

\newlength{\colwidth}
\setlength{\colwidth}{\columnwidth}

\addtolength\voffset{0.8cm}

%% LLT: Adding these manually so that they're already defined
%% during the very first compile
%% Need each line for each table
\makeatletter
\def\deluxe@table@width@i{0pt}
\def\deluxe@table@width@ii{0pt}
\def\deluxe@table@width@iii{0pt}
\makeatother

\newcommand{\ds}{\displaystyle}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\bee}{\begin{eqnarray}}
\newcommand{\eee}{\end{eqnarray}}

\newcommand{\ov}{ \Omega_{\rm \Lambda} }
\newcommand{\om}{ \Omega_{\rm m} }
\newcommand{\mob}{ M_{\rm obj} }
\newcommand{\mstar}{ M_{*}(a) }
\newcommand{\msol}{\hbox{${\rm M}_\odot$}}

\newcommand{\rvir}{\hbox{$r_{\rm 200}$}}
\newcommand{\mvir}{\hbox{$M_{\rm vir}$}}
\newcommand{\Mvir}{\hbox{$M_{\rm vir}$}}
\newcommand{\vvir}{\hbox{$v_{\rm 200}$}}
\newcommand{\tvir}{T_{\rm vir}}
\newcommand{\mthresh}{M_{\rm t}}
\newcommand{\nsats}{n_{\rm sats}}
\newcommand{\vmax}{{\rm v}_{\rm max}}
\newcommand{\zreion}{z_{\rm reion}}
\newcommand{\LCDM}{$\Lambda$CDM }
\newcommand{\sigate}{ \sigma_{8}}
\newcommand{\actoy}{a_{c,{\rm toy}}}
\newcommand{\K}{{\rm K}}

\newcommand{\DME}{D_{\rm M31}}
\newcommand{\vMEI}{v_{\rm M31}}
\newcommand{\MMEI}{M_{\rm M31}}
\newcommand{\DMEE}{D_{\rm M33}}
\newcommand{\vMEE}{v_{\rm M33}}
\newcommand{\MMEE}{M_{\rm M33}}
\newcommand{\DLMC}{D_{\rm LMC}}
\newcommand{\vLMC}{v_{\rm LMC}}
\newcommand{\MLMC}{M_{\rm LMC}}
\newcommand{\MMW}{{\rm M}_{\rm MW}}
\newcommand{\Nsubs}{N_{\rm subs}}

\newcommand{\LikePR}{\textbf{Pr(d$\vert$x)}}

\def\rx{X}
\def\ry{Y}
\def\rz{Z}
\def\vx{v_X}
\def\vy{v_Y}
\def\vz{v_Z}
\def\distance{D}
\def\vrad{v_{\rm rad}}
\def\vtan{v_{\rm tan}}

\def\pr{{\rm Pr}}
\def\pars{\mathbf{x}}
\def\parsi{x_i}
\def\data{\mathbf{d}}
\def\datai{d_i}
\def\datap{\mathbf{d}^{\rm p}(\pars)}
\def\datapi{d^{\rm p}_{i}(\pars)}

\newcommand{\Msun}{{\rm M}_\odot}
\newcommand{\hinv}{h^{-1}}
\newcommand{\Mpc}{{\rm Mpc}}
\newcommand{\hmpc}{\hinv\Mpc}
\newcommand{\kpc}{{\rm kpc}}
\newcommand{\kms}{{\rm km s}^{-1}}

\newcommand{\rhobar}{\langle \rho \rangle}
\newcommand{\sig}{ \langle \sigma v \rangle }
\newcommand{\sigwig}{ {\widetilde \sigma} }
\newcommand{\rbh}{ r_{\rm bh} }
\newcommand{\rmax}{ r_{\rm max} }
\newcommand{\kmsmpc}{\, \rm{km}\,  \rm{s}^{-1}\, \rm{Mpc}^{-1}}
\newcommand{\lsim}{\lower.5ex\hbox{\ltsima}}
\newcommand{\gsim}{\lower.5ex\hbox{\gtsima}}
\newcommand{\ltsima}{$\; \buildrel < \over \sim \;$}
\newcommand{\gtsima}{$\; \buildrel > \over \sim \;$}
\def \etal      {\hbox{et al.} }

\newcommand{\bolshoi}{{\sc Bolshoi }}
\newcommand{\consuelo}{{\sc Consuelo }}
\newcommand{\subfind}{\texttt{\footnotesize SUBFIND }}
\newcommand{\CMBFAST}{\texttt{\footnotesize CMBFAST }}

\def\Sref#1{Section~\ref{#1}}
\def\Fref#1{Figure~\ref{#1}}
\def\Tref#1{Table~\ref{#1}}
\def\Eref#1{Equation~\ref{#1}}
\def\Aref#1{Appendix~\ref{#1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% RESULTS:

\def\MLG{M_{\rm LG}}
\def\MMW{M_{\rm MW}}
\def\MEI{M_{\rm M31}}
\def\MEE{M_{\rm M33}}

\def\CMEI{c_{\rm M31}}
\def\CMW{c_{\rm MW}}

% Local group mass: MW+M31 from M31 kinematics:
\def\MPAIRestimate{2.40}
\def\MPAIRerrorplus{1.77}
\def\MPAIRerrorminus{1.20}

\def\MPAIREXestimate{2.88}
\def\MPAIREXerrorplus{2.13}
\def\MPAIREXerrorminus{1.47}

% Local group timing argument quantities:
\def\ATA{A_{200}}
\def\MTA{M_{LG,TA}}

% Mass of M31+M33 system, from M33 kinematics:
\def\MbAestimate{xx}
\def\MbAerrorplus{xx}
\def\MbAerrorminus{xx}

% Individual halo concentration estimates for M31 and MW

\def\CMWestimate{8.53}
\def\CMWerrorplus{1.89}
\def\CMWerrorminus{2.59}

\def\CMEIestimate{8.61}
\def\CMEIerrorplus{1.69}
\def\CMEIerrorminus{2.32}

% Individual galaxyhalo mass estimates from both M31 and M33 kinematics:
\def\MMWestimate{0.87}
\def\MMWerrorplus{0.30}
\def\MMWerrorminus{0.30}

\def\MEIestimate{1.29}
\def\MEIerrorplus{0.71}
\def\MEIerrorminus{0.40}

\def\MEEestimate{0.17}
\def\MEEerrorplus{0.16}
\def\MEEerrorminus{0.10}

\def\MLMCestimate{0.13}
\def\MLMCerrorplus{0.08}
\def\MLMCerrorminus{0.06}

\def\MLGestimate{2.09}
\def\MLGerrorplus{0.93}
\def\MLGerrorminus{0.61}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage[usenames]{color}
\newcommand{\risa}[1]{\textcolor{red}{\bf #1}}
\newcommand{\phil}[1]{\textcolor{ForestGreen}{\bf #1}}
\newcommand{\marc}[1]{\textcolor{blue}{\bf #1}}
\newcommand{\question}[1]{\textcolor{blue}{\bf #1}}
\newcommand{\todo}[2]{{\bf To do (#1): #2}}
\newcommand{\new}[1]{{\bf #1}}

\newenvironment{inlinefigure}{\def\@captype{figure}
\noindent\begin{minipage}{0.999\linewidth}\begin{center}}
{\end{center}\end{minipage}\smallskip}

\shortauthors{Williamson et al.}
\shorttitle{The Mass of the Local Group}

% ----------------------------------------------------------------------------

\begin{document}

\title{The Dark Matter Distribution of the Local Group inferred from
Cosmological Simulations: I. The Halo Masses of M31, M33 and the Milky Way}

\author{Marc Williamson}
\author{Philip J.~Marshall}
%\author{Michael T. Busha\altaffilmark{1,2}} 
\author{Yao-Yuan Mao} 
\author{Risa H.~Wechsler}
\affil{Kavli Institute for Particle Astrophysics and Cosmology
  \& Physics Department, Stanford University, Stanford, CA 94305, USA}
\affil{SLAC National Accelerator Laboratory, Menlo Park, CA, 94025, USA}
%\altaffiltext{1}{Kavli Institute for Particle Astrophysics and Cosmology, Stanford, CA, 94305, USA
%{\tt mwillia1@stanford.edu, pjm@slac.stanford.edu, rwechsler@stanford.edu}}
%\altaffiltext{2}{Elastica}

% ----------------------------------------------------------------------------

\begin{abstract}

We study the Local Group of galaxies by identifying analogous systems in a large suite of N-body simulations.  The halo catalogs are treated as a large sample drawn from the cosmological prior PDF, and Local Group analogs are identified by matching the observed kinematics of the system.  Weights are calculated for the  halo sample by finding the likelihoods for the simulated kinematic properties compared to the observed galactocentric distances and relative velocities between the Milky Way, M31, M33, and LMC.  We present a novel method for resolving small number statistics issues: namely, the simulation prior can be fit with a Gaussian Mixture Model, allowing arbitrarily large samples to be drawn. We study the effect of including the existence and kinematics of M33 and LMC on the posterior distributions of the local group mass by using three types of samples: from only the "observed" galactocentric distance, radial velocity, and tangential velocity of M31, we infer the virial mass of the local group to be $\MLG =
(\MMW + \MEI) = (\MPAIRestimate^{+\MPAIRerrorplus}_{-\MPAIRerrorminus}) \times
10^{12} \Msun$. Including only the existence of M33 and LMC as a condition for identifying local group analogs results in a different inference of the mass, $\MLG =
(\MMW + \MEI) = (\MPAIREXestimate^{+\MPAIREXerrorplus}_{-\MPAIREXerrorminus}) \times
10^{12} \Msun$. Finally including the kinematics of M33 and LMC we infer the mass of the local group to be $\MLG =
(\MMW + \MEI) = (\MLGestimate^{+\MLGerrorplus}_{-\MLGerrorminus}) \times
10^{12} \Msun$.  We are also able to estimate the halo masses of the four galaxies
individually, finding these to be $\MEI =
(\MEIestimate^{+\MEIerrorplus}_{-\MEIerrorminus}) \times 10^{12} \Msun$, $\MEE
= (\MEEestimate^{+\MEEerrorplus}_{-\MEEerrorminus}) \times 10^{12} \Msun$, 
and $\MMW = (\MMWestimate^{+\MMWerrorplus}_{-\MMWerrorminus}) \times 10^{12}
\Msun$, $\MLMC = (\MLMCestimate^{+\MLMCerrorplus}_{-\MLMCerrorminus}) \times 10^{12}
\Msun$.  Our Milky Way mass is lower than previous estimates.
\end{abstract}

\keywords{%
Galaxy: halo, fundamental parameters -- 
galaxies: haloes, fundamental parameters -- 
galaxies: Local Group -- 
galaxies: individual: M31, M33 --
cosmology: dark matter}

% ----------------------------------------------------------------------------

\section{Introduction}
\label{sec:intro}

The Local Group (LG) of galaxies consists of two large spiral galaxies,
Andromeda (M31) and the Milky Way (MW), and a population of smaller
satellite galaxies, the biggest of which are Triangulum (M33) and the Large Magellanic Cloud (LMC). We know
relatively little about the distribution of dark matter throughout the Local
Group volume, beyond simple estimates of the total mass of the group
estimated from its members kinematics (e.g. \citealt{VdM08}, hereafter vdMG08).
Our knowledge of the dark matter halos of the group member galaxies is
similarly sparse: virial masses have been estimated from extrapolations of
simple models fitted to stellar or gas kinematics  \citep[e.g.][]{xue2008} or to the observed positions and velocities of satellite galaxies \citep[e.g.][]{watkins2010}.

Our goal is to tightly constrain the mass of the Local Group by combining observational data on the  positions and velocities of the four largest LG members with our understanding of structure formation
in $\Lambda$CDM universes, as detailed in halo catalogs extracted from
cosmological N-body simulations. The halo catalogs are treated as a set of samples
drawn from a very informative, highly correlated prior PDF for halo properties. 
We apply broad isolation criteria based on the observed distances between the MW, M31, M33, and LMC, 
in addition to a local environment isolation criterion in order to identify a sample of LG analogs in the catalog.
Each LG analog is then weighted by the likelihood that it is a Local Group analog, 
calculated by comparing the kinematics of the simulated system to the observed galactocentric distances and relative velocities between the MW, M31, M33, and LMC.
 This method makes natural physical assumptions about the structure of the LG and avoids difficult to interpret stellar kinematics data.  In \cite{busha2011mass} (hereafter B11) we presented a new
measurement of the Milky Way halo mass, made in this way using the
\bolshoi simulation of \citet{Bolshoi} and the distances to and speeds
of the Magellanic Clouds. In this work, we use the lower resolution but
higher volume \consuelo simulations, and constrain them with the radial
velocities, proper motions and distances of M31, M33, and LMC.

This research extends the work of \cite{gonzalez2014mass} by considering the effects of including M33 and LMC in the definition of a Local Group analog. Our work also provides estimates of the masses for all three individual halos that are part of the Local Group, MW, M31, M33, and LMC. In a subsequent
paper we will explore other properties of these halos. As shown by Gonzalez et al, the Bayesian analysis used in this work is more robust than the Timing Argument of \cite{Timing} which relies on simplifying assumptions about the structure of halos and the evolution of their positions.

In short, we seek answers to the following questions:
\begin{itemize}
\item What is the mass of dark matter in the Local Group?
\item How do the existence and kinematics of M33 and LMC impact the Local Group mass?
\item What are the masses of the dark matter halos of M31 and the Milky
Way?
\item How sensitive is this method to the isolation criteria used to define an LG-analog?
\end{itemize}

This paper is structured as follows.  In 
\Sref{sec:method} 
we review the
halo inference methodology, and introduce the \consuelo N-body
simulation suite we use. In 
\Sref{sec:coord} 
we review the coordinate systems and transformations used in this work. In 
\Sref{sec:results} 
we present
our results on the mass of the Local Group and its members, and then
discuss those results in the context of previous work in
\Sref{sec:discuss}. 
We conclude in 
\Sref{sec:conclude}.

Throughout this work we assume a \LCDM cosmology, with $\Omega_{\rm m} =
0.25$, $\Omega_{\rm \Lambda} = 0.75$, and $H_0 = 70 \kms \Mpc^{-1}$. Where
a point estimate of a parameter is given, its value is the median of the
one-dimensional marginalised posterior probabilty distribution, and its
quoted uncertainty describes the 68\% credible region.
We define a halo's mass by its $M_{vir}$, the mass enclosed within the virial radius, $r_{vir}$,
at the redshift under consideration.

%-----------------------------------------------------------------------------
% ----------------------------------------------------------------------------

\section{Methodology}
\label{sec:method}

In this section we review the halo inference introduced in B11, and describe
the \consuelo cosmological N-body simulations that we use in this work.

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\subsection{Halo catalogs: samples from the halo prior PDF}
\label{sec:sampling}

The halos in the \consuelo catalogs are characterized by a set of parameters, \textbf{x}, which includes the halo positions, velocities, mass, angular momentum, etc.  The \consuelo halo catalog is therefore a large sample drawn from the underlying PDF for halo characteristics, \textbf{Pr(x)}.  By weighting the \consuelo halos by how much they resemble our own Local Group, we can infer the properties of the Local Group and its members.  Thus given our observational data, \textbf{d}, of the LG, we are interested in the posterior distributions of the LG parameters.
\begin{center}

\textbf{Pr(x$\vert$d) $\propto$ Pr(d$\vert$x)Pr(x)}

\end{center}
The term \textbf{Pr(x)} is the cosmological prior, henceforth referred to as the \consuelo prior, and the weighting factor, \textbf{Pr(d$\vert$x)} is the likelihood function for the parameters given our data. In section
\ref{sec:gmm_gof_app}
we discuss our continuous approximation of the likelihood function.
%\end{}

\subsection{Importance Sampling}
\label{sec:importance}
This work follows the statistical methodology of B11. First, we give a brief background on importance sampling before applying it to our situation. A deeper introduction is given by \cite{imp_book}. Let \textbf{P(x)} be a distribution from which we want to draw a sample, but cannot due to the complex nature of \textbf{P(x)}. Assume though, that there is another distribution, \textbf{P'(x)} that we can evaluate densities for, such that $\textbf{P(x)}=\frac{\textbf{P'(x)}}{Z_{P}}$. Now consider a simpler distribution, \textbf{Q(x)}, from which we can draw a samples and evaluate densities to within a multiplicative factor. Suppose that we wish to evaluate the mean of some parameter, $\phi(x)$, whose true prior PDF is \textbf{P(x)}. We cannot simply use the sample drawn from \textbf{Q(x)} because \textbf{Q(x)} will overestimate and underestimate \textbf{P(x)} for certain values of \textbf{x}. This problem is addressed by weighting each sample from \textbf{Q(x)} by the ratio 
\begin{center}
$w = \frac{\textbf{P'(x)}}{\textbf{Q'(x)}}$
\end{center}

\par
For our application, the complex PDF from which we would ideally be able to draw sample Local Groups is \textbf{P(x$\vert$d)}. The associated distribution that we can evaluate to within a constant factor is given by Bayes Theorem, shown at the end of section 
\ref{sec:sampling}. 
If it were possible to draw samples directly from the conditional PDF \textbf{P($x_{i}\vert$ d)}, then it would be easy to calculate the mean values of LG parameters like mass. However, we can only draw samples from the \consuelo catalog.  Therefore, in order to infer the masses of the LG and its member halos, we must weight the samples drawn from the prior by the factor:

\begin{equation}
\begin{aligned}
w &= \frac{\textbf{Pr(d$\vert$x)Pr(x)}}{\textbf{Pr(x)}} \\
 &= \textbf{Pr(d$\vert$x)}
\end{aligned}
\end{equation}
which is simply the likelihood. Therefore, we draw samples from the prior, \textbf{Pr(x)} and weight them according to the likelihood function in order to approximate drawing a sample from the conditional PDF. In order to evaluate the likelihood for analog LG systems found in the \consuelo catalogs, we require a continuous model of the likelihood function.  In addition, the observed data is based in an Equatorial reference frame, which is incompatible with the Galactocentric reference frames used in the \consuelo simulation.  Thus there are three steps necessary to be able to weight the \consuelo LG analogs by their likelihood, \LikePR.  
\begin{enumerate}
\item Transform observed data from Table 
\ref{table:obs}, 
\textbf{$d$}, to M31 galactocentric reference frame.
\item Calculate observed values of the likelihood parameters, $(D, v_{r}, v_{t})$ for M31, M33, and LMC.
\item Fit the observed distribution of the likelihood parameters with a GMM.
\end{enumerate}
The coordinate systems and transformations used in this work are detailed in section 
\ref{sec:coord}.

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\subsection{The \consuelo Simulation}
\label{sec:sampling_b}

As mentioned above, we use the \consuelo set of simulations, run as part of the Large Suite of Dark Matter Simulations (LasDamas). \consuelo is a collision-less, dark matter only simulation tracking $1400^{3}$ particles, each with individual mass $1.87\times10^{9} \Msun /h$. The cosmology is a flat $\Lambda$CDM with parameters $\Omega_{M}=0.25$, $\Omega_{\Lambda}=0.75$, $h=0.7$, $n_{s}=1.0$, and $\sigma_{8}=0.8$.  
We apply our Local Group isolation criteria to the redshift zero snapshot of 17 \consuelo boxes.
This yields approximately $200000$ \consuelo Local Group analogs that we use for constructing our prior PDF for halo characteristics.

%\todo{Marc}{Add number of halos, vir vs 200 for mass and concentration.}

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\subsection{The Local Group prior}
\label{sec:lg_prior}

In this section we present our methods for drawing samples from the \consuelo prior, $\textbf{Pr(x)}$.  We work with two general definitions of a Local Group analog:
\begin{itemize}
\item Pair -- A system with a MW analog and an M31 analog.
\item Quad -- A system with a MW analog and an M31 analog.  The M31 analog has an associated M33 analog, and the MW analog has an associated LMC analog.
\end{itemize}

\subsubsection{Isolation Criteria}
\label{sec:iso_crit}

We assume that the MW and M31 are host halos (as opposed to satellites), while no assumption is made about M33 or LMC. In addition, we make a cut at $v_{max} > 80$ km/s to ensure that the halos are well resolved. 
\Fref{fig:iso_diagram} shows a 2D schematic of a sample Quad LG analog.
The isolation criteria used to identify LG analogs in the catalogs are listed below:

\begin{figure}[h]
\includegraphics[width=\linewidth]{figures/isd.png}
\caption{2D example of a Quad LG analog. $D_{pair} < 1$ Mpc.  $\DMEE, \DLMC < 0.4$ Mpc.  $D_{iso} < 2$ Mpc.}
\label{fig:iso_diagram}
\end{figure}

\begin{itemize}
\item The distance between the MW and M31 analogs is less than 1 Mpc.  This is ~5 sigma larger than the true value, $D=0.77$ Mpc.
\item Within 2 Mpc of the MW-M31 midpoint, the MW and M31 are the two largest halos.
\item For a given M31 (MW) analog, the associated M33 (LMC) is the most massive halo within 0.4 Mpc.
\end{itemize}
The second criterion enforces that neither MW nor M31 are within the virial radius of a much larger halo.  This is justified by the fact that the Local Group is not part of a larger galaxy cluster.

\todo{Marc}{Comment on labeling schemes for halos?}

\subsection{Gaussian Mixture Models}
\label{sec:gmm}

We use Gaussian Mixture Models to fit two different distributions in this work: once to fit the likelihood function (henceforth referred to as the LGMM), $\textbf{Pr(d$\vert$x)}$, and once to fit the \consuelo prior (henceforth referred to as the PGMM), $\textbf{Pr(x)}$. 
The GMM fit to the \consuelo prior is presented as an alternative to using the \consuelo halo catalog to approximate the cosmological PDF for halo characteristics. This is a novel approach that allows arbitrarily large fake halo catalogs to be generated, assuming the GMM fits the \consuelo prior well. Details on the GMM method and goodness of fit are presented in 
\Sref{sec:gmm_gof_app}

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\subsection{Observational Constraints}
\label{sec:data}

The observational data used to constrain our two and three-halo cosmological models of the Local Group is shown in Table \ref{table:obs}.

\def\arraystretch{1.4}
\capstartfalse
\begin{deluxetable*}{ccccc}[h]
\tabletypesize{\footnotesize}
\tablecaption{\label{table:obs}
\tablewidth{500pt}
Observables of MW, M31, and M33.}
\startdata
\hline
\hline
                     & M31               & M33         &LMC               & Reference          \\
\hline

$\alpha$ / deg          & $ 10.68 \pm 0.08$ & $ 23.46 \pm 0.08$ &$78.76 \pm 0.52$& NED   \\
$\delta$ / deg          & $ 41.27 \pm 0.08$ & $ 30.66 \pm 0.08$ &$-69.19 \pm 0.25$& NED   \\
$D$ / Mpc               & $  0.77 \pm 0.04$ & $ 0.79 \pm 0.23$ &$0.0500 \pm 0.0025$& vdM12, vdM02LMC   \\
$v_{r}$ / km s$^{-1}$  & $ -301.8 \pm 1.00$ & $ 180.0 \pm 1.00$ &$262.0 \pm 3.40$& vdMG08, vdM02LMC   \\
$v_{W}$ / km s$^{-1}$  & $  -125.2 \pm 30.8$ &  N/A & N/A & vdM12   \\
$v_{N}$ / km s$^{-1}$  & $  -73.8 \pm 28.40$ &   N/A & N/A & vdM12   \\
$\mu_{W}$ / $\mu$as yr$^{-1}$ & $  -42.2 \pm 12.30$ & $ 4.70 \pm 3.20$ &$-1910.00 \pm 160.00$& Sohn12, vdMG08, K13   \\
$\mu_{N}$ / $\mu$as yr$^{-1}$ & $  -30.9 \pm 11.7$ & $ -14.1 \pm 6.4$ &$229.00 \pm 160.00$& Sohn12, vdMG08, K13   \\

\hline
\enddata
\tablecomments{
The first two quantities, right ascension and declination, are measured so precisely that the uncertainty can be ignored. $(\mu_{W}, \mu_{N})$ are the proper motions in spherical coordinates, while $(v_{W}, v_{N})$ are the associated linear velocities. The $(\delta v_{rot})$ quantities are a correction for the internal rotation of the galaxies being observed (M31 and M33). This correction is only necessary for calculating the $(v_{W},v_{N})_{M33}$ velocities for M33.  The velocities for M31 are obtained from the literature.
}
\medskip
\end{deluxetable*}
\capstarttrue

As discussed in the section on coordinate transformations, this data is measured in Equatorial coordinates. In order to obtain the parameters used to define a Local Group, $(D, v_{r}, v_{t})_{MW,M33}$, we apply the described coordinate transformations to change to the M31 galactocentric reference frame.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

% ----------------------------------------------------------------------------
\def\arraystretch{1.4}
\capstartfalse
\begin{deluxetable*}{lccccccc}[h!]
\tabletypesize{\footnotesize}
\tablecaption{\label{table:mass_res}
Log Posterior Masses $\log(M_{vir}/\Msun)$}
\tablewidth{500pt}
\startdata
\hline
\hline
                      & MW               & M31           &M33           & LMC      & LG           &N     &N95\\
\hline 
Quad (GMM prior)   &  $11.94^{+0.13}_{-0.18}$ & $12.11^{+0.19}_{-0.16}$ & $11.23^{+0.29}_{-0.37}$& $11.10^{+0.21}_{-0.25}$ & $12.32^{+0.16}_{-0.15}$ & $4\times 10^{8}$ & $721$\\
Pair (M33, LMC Existence)   &  $11.89^{+0.30}_{-0.32}$ & $12.30^{+0.26}_{-0.34}$ & $11.16^{+0.32}_{-0.27}$ & $11.14^{+0.31}_{-0.26}$&$12.46^{+0.24}_{-0.31}$ &$1\times 10^{7}$ &$8.87\times 10^{4}$\\
Pair (GMM prior)   &  $11.72^{+0.35}_{-0.40}$ & $12.23^{+0.27}_{-0.33}$ & N/A & N/A & $12.38^{+0.24}_{-0.30}$ &$1\times 10^{7}$ &$5.05\times 10^{4}$  \\
\hline

\enddata
\tablecomments{
The first column lists the type of isolation criteria used to define a Local Group analog. Quad and Pair are as defined in the text, and M33, LMC Existence indicates that only the existence, not the dynamics, are considered in calculating the Likelihood. $N$ is the unweighted sample size, while $N95$ is the number of LG analogs contributing to $95\%$ of the cumulative likelihood.
}
\medskip
\end{deluxetable*}
\capstarttrue

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
\label{sec:results}

In this section, we present the results for our mass posterior calculations. Our comprehensive results for each type of isolation criteria used in this work are shown in Table \ref{table:mass_res}. 
\Fref{fig:LvsPGMM} shows which regions of the \consuelo prior distribution are significantly weighted by the Likelihood function.

\begin{figure*}
\includegraphics[width=\linewidth]{figures/LvsPGMM.pdf}
\caption{The Quad likelihood contours (red) show which regions of the \consuelo prior are strongly weighted. The LMC uncertainties in proper motion and distance are increased by a factor of four to increase the number of highly weighted Quad systems.}
\label{fig:LvsPGMM}
\end{figure*}

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\subsection{LG and Halo Masses}
\label{sec:nums}
\Fref{fig:Q_GMMP_all_Mvir} displays the posterior distributions for each of the halo masses in the Quad system. 
The lower mass tails of the LMC and M33 1D histograms show that the halos are resolved in the \consuelo simulations. 
As expected, $\MEI$ correlates strongly with $\MLG$ and $\MMW$.
We find $\MMW = (\MMWestimate^{+\MMWerrorplus}_{-\MMWerrorminus}) \times 10^{12}
\Msun$, $\MLMC = (\MLMCestimate^{+\MLMCerrorplus}_{-\MLMCerrorminus}) \times 10^{12}
\Msun$, $\MEI =
(\MEIestimate^{+\MEIerrorplus}_{-\MEIerrorminus}) \times 10^{12} \Msun$, $\MEE
= (\MEEestimate^{+\MEEerrorplus}_{-\MEEerrorminus}) \times 10^{12} \Msun$, 
and $\MLG = (\MLGestimate^{+\MLGerrorplus}_{-\MLGerrorminus}) \times 10^{12}
\Msun$.
Our estimate for $\MMW$ supports a low Milky Way mass, and our estimate for $\MLMC$ supports a large LMC mass.

\begin{figure*}[ht]
\includegraphics[width=\linewidth]{figures/Q_GMMP_all_Mvir.pdf}
\caption{Posterior distributions of log virial mass for MW, M31, M33, LMC, and the LG, where the LG mass is the sum of MW and M31 masses.}
\label{fig:Q_GMMP_all_Mvir}
\end{figure*}

%\begin{itemize}
%\item Present MW, M31, M33, and LMC masses.
%\item Relative size of MW and M31.
%\item Are LMC and M33 masses resolved? (Are 1D histograms cut off)
%\item Present M200 masses?
%\end{itemize}

\subsection{Halo Concentrations}
\label{sec:conc}
\Fref{fig:Q_GMMP_M31_MvsC} and \Fref{fig:Q_GMMP_MW_MvsC} display the posterior predictions for the concentration-mass relation for M31 and the MW respectively. 
We find 
$\CMW = \CMWestimate^{+\CMWerrorplus}_{-\CMWerrorminus}$ and 
$\CMEI = \CMEIestimate^{+\CMEIerrorplus}_{-\CMEIerrorminus}$. 
The plots also show the prior concentration-mass relation in the relevant mass regime. We note that the posterior 2D distributions show that there is no correlation between mass and concentration for both M31 and the MW. 
This is somewhat surprising, and it indicates that M31 and the MW can have more extreme concentrations for their mass than would be expected from simply analyzing the \consuelo prior.

\begin{figure}[ht]
\includegraphics[width=\linewidth]{figures/Q_GMMP_M31_MvsC.pdf}
\caption{Posterior distributions of log mass and concentration for M31. The concentration-mass relation with an uncertainty of $1\sigma$ for this mass regime is shown overlayed.}
\label{fig:Q_GMMP_M31_MvsC}
\end{figure}

\begin{figure}[ht]
\includegraphics[width=\linewidth]{figures/Q_GMMP_MW_MvsC.pdf}
\caption{Posterior distributions of log mass and concentration for MW. The concentration-mass relation with an uncertainty of $1\sigma$ for this mass regime is shown overlayed.}
\label{fig:Q_GMMP_MW_MvsC}
\end{figure}

%\begin{itemize}
%\item Present MW, M31, M33, and LMC %concentrations.
%\item Mvir-Cvir correlation?
%\item Relative Cvir of MW and M31.
%\end{itemize}

\subsection{The Effect of M33 and LMC}
\label{sec:existence}
\Tref{table:mass_res} shows that the posterior mass predictions are quite sensitive to the existence and kinematics of M33 and the LMC. 
Ignoring both satellites as was done in \cite{gonzalez2014mass} underestimates $\MMW$ and overestimates $\MEI$. These effects almost negate each other for the Local Group mass. \Fref{fig:ta_plot} compares our results to those of the Timing Argument \citep{Timing}. When the kinematics of M33 and the LMC are included as constraints on the LG definition, the TA significantly overestimates the LG mass.
Therefore, the classic two body approach of the Timing Argument \citep{Timing}, even with the modifications to include local environment variables in the Local Group definition from \cite{gonzalez2014mass}, is insufficient to capture the complete behavior of the Local Group.

\begin{figure}[ht]
\includegraphics[width=\linewidth]{figures/timing_plot.pdf}
\caption{Ratio of the true MW-M31 pair mass to Timing Argument mass estimate. When only the kinematics of the MW-M31 system are used as constraints, the median of $A_{vir}$ is close to unity. However, when the kinematics of M33 and the LMC are included, the Timing Argument overestimates the LG mass, driving the median of $A_{vir}$ lower.}
\label{fig:ta_plot}
\end{figure}

% ----------------------------------------------------------------------------

\section{Discussion}
\label{sec:discuss}

\subsection{Comparison to the Timing Argument}
\label{sec:TA}

\begin{itemize}
\item Brief summary of TA: Eqs. 1-3 from LW08
\item What did Li and White calculate, and what were their results?
	\subitem LG Mass
	\subitem Bias of the TA
\item Drawbacks of TA
	\subitem Ignores mass evolution
	\subitem Ignores tangential velocity
	\subitem Ignores environment
\item How did Gonzalez et al. update the TA? 
	\subitem Tangential velocity lowers $A_{vir}$.
	\subitem Local density has negligible effect.
\item What do we do differently?
	\subitem Consider the kinematics and existence of M33 and LMC.
	\subitem Make sure to separate the effects of adding LMC and M33 and adding the tangential velocities.
	\subitem Which kinematics are the most important?
\item Figure 6
\end{itemize}

\subsection{Mass of the Milky Way}
\label{sec:MW_discuss}
\todo{Marc}{See what the updated result is for $\MMW$. This will determine what goes in this section.}
\begin{itemize}
\item Brief summary of previous methods. See intro.
\item What do we do differently?
	\subitem M33, LMC
	\subitem compare our results to literature.
	\subitem \subitem Specifically mention differences (and possible problems) with similar methods: Busha, Gonzalez, Kafle, Penarrubia.
\item Consequences for missing satellite problem?
\end{itemize}

\subsection{Mass of LMC}
\label{sec:LMC_discuss}
\todo{Marc}{Update $\MLMC$ value.}
\todo{Marc}{$\MLMC$ lower bound vs $v_{res}$ cut.}
\begin{itemize}
	\item Brief summary of previous models for mass ratio of LMC/MW (see Pena. for sources)
	\item Note use of GMM prior allows us to have enough highly weighted LMC's to get good stats.
	\item Present our calculation of $f_{LMC/MW}$.
		\subitem How does it compare to lit?
		\subitem Consequences of being large?
	\item Is our LMC mass impacted by resolution? By how much?
		\subitem New Figure.
\end{itemize}

\subsection{GMM Prior Goodness of Fit}
\label{sec:gmm_discuss}
\todo{Marc}{Discussion or Appendix?}
\begin{itemize}
	\item What do we use GMM prior for?
		\subitem Why is it important that it must fit distribution well?
	\item GMM prior 1D and 2D marginalizations.
	\item What do we need to check?
		\subitem How much volume (num of consuelo boxes) do we need? New Figure: $\MMW$ post vs GMM volume.
		\subitem How many components in the GMM? New Figure: $\MMW$ post vs num of GMM components.
	\item Pair posterior estimates using GMM prior vs Pair posterior estimates using $\consuelo$.
	\item Implications of good GMM prior.
		\subitem Larger samples
		\subitem Ability to use stronger constraints.

\end{itemize}

Using the Quad definition of the LG, we find that 
$\MEI =
(\MEIestimate^{+\MEIerrorplus}_{-\MEIerrorminus}) \times 10^{12} \Msun$, $\MEE
= (\MEEestimate^{+\MEEerrorplus}_{-\MEEerrorminus}) \times 10^{12} \Msun$,
both of which are consistent at the $1\sigma$ level with previous results \citep{penarrubia2015timing, m33_mass}.
Our posterior constraint on the Milky Way mass, 
$\MMW =
(\MMWestimate^{+\MMWerrorplus}_{-\MMWerrorminus}) \times 10^{12} \Msun$,
is smaller than most previous studies, although still consistent at the $1\sigma$ level with $\MMW = 1.0\times10^{12} \Msun$. 
Since the number of subhalos at a given mass scales with host halo mass \citep{wang2012missing}, a low Milky Way mass is one possible solution to the "Missing Satellite Problem."  Comparing to Figure 5 in Wang et al, we see that there is about a 60 percent chance for a host halo of our mass $\MMW$ to have three or fewer subhalos with $V_{max} > 30$ km/s. 
\par
In addition to finding that the MW is less massive than previously thought, our method shows that the LMC is much larger relative to the MW than expected.  
We calculate the ratio $f_{c} = \MLMC/\MMW$ to be $f_{c} = 0.13^{+0.13 (0.52)}_{-0.05 (0.09)}$, ruling out models where $f_{c} < 0.04$ at the 95 percent level.
Such a large LMC mass relative to the Milky Way could explain the systematic mass overestimate of the Timing Argument seen in \Fref{fig:ta_plot}.
\par
The posterior mass and concentration  calculations presented in this paper depend on the goodness of fit of the PGMM and LGMM, as discussed in \Sref{sec:gmm_gof_app}. 
We find that variations in the number of components in either GMM do not alter the posterior estimates.
As discussed by \cite{kafle2014shoulders}, the MW-LMC point mass approximation affects the posterior calculations by at most 5 percent.
The uncertainties reported in this paper are purely statistical errors.
\par
The work in this paper can easily be extended in numerous ways. 
We have shown that a simple application of unsupervised machine learning (GMM's) can solve small number statistics problems often encountered in working with cosmological simulations.
Other types of learning algorithms (clustering, regression, etc) should be explored.
In this paper, we focus on studying halo mass and concentration. 
There are many other halo properties that could give valuable insight into evolution of the Milky Way and Local Group.
Given how large the large mass of the LMC compared to the MW, it would be interesting to know about the pair's relative orientation and angular momentum.

\section{Summary}
\label{sec:conclude}
\todo{Marc}{Update Summary as results are updated.}
\par
This paper uses the galactocentric kinematic properties of the Local Group to infer the masses of the MW, LMC, M31, and M33. 
We apply a Bayesian method to model the likelihood of a simulated system of halos being a Local Group and use the likelihood to weight a sample of LG analogs drawn from the \consuelo prior. 
This allows us to derive posterior PDF's for any halo property tracked by the simulation, most notably halo mass and concentration. 
We present a novel use of Gaussian Mixture models to approximate the prior halo PDF, allowing us to surpass current computational limits on the number of halos that it is possible to simulate.
The GMM prior approximation enables the use of M33 and the LMC's kinematics as constraints on the definition of the Local Group for the first time.
\acknowledgments{Acknowledgments go here.}

\bibliographystyle{apj}
\bibliography{references}
\clearpage

\appendix
\section{Coordinate Transformations}
\label{sec:coord}
In this section we describe the coordinate systems used in this paper and present the transformations between them.  As mentioned in the introduction, this work uses the galactocentric dynamics of the MW-M31-M33-LMC system to define a LG analog. Specifically, we use the parameters $\distance$, $\vrad$, and $\vtan$ of M31 and LMC measured in the MW galactocentric frame, and the same parameters for M33 measured in the M31 galactocentric frame. 
The observed values of properties like position, distance, and proper motion for these galaxies are made in equatorial coordinates.  However, the corresponding properties in the \consuelo catalogs are measured in galactocentric coordinates.  Therefore, it is necessary to convert the observations of M31 and LMC to the MW galactocentric reference frame, and the observations of M33 to the M31 galactocentric reference frame.  The rest of this section defines the intermediate coordinate systems used in this work and the coordinate transformations between them.

\subsection{Coordinate System Definitions}
\label{sec:c_sys_defs}
Three types of coordinate systems are used in this work: Geocentric Equatorial, Heliocentric Galactic, and Galactocentric.  Before we describe the mathematical transformations between these systems, it is important to thoroughly understand the definitions of the systems. 
\par The equatorial system is centered on the Earth and based on the concept of a celestial sphere (ie all stars are projected onto a sphere of infinite radius centered on the Earth). Two angles characterize the positions of objects on the celestial sphere: right ascension and declination. Right ascension, $\alpha$, is the angle measured eastwards along the celestial equator between the vernal equinox and the hour circle passing through an object.  Declination, $\delta$, is the angular distance of an object above or below the celestial equator.

\begin{figure}[ht]
\includegraphics[width=\linewidth]{figures/eq_sys.png}
\caption{Illustration of the equatorial coordinate system. The origin is the center of the Earth. The $X$ axis points towards the Vernal Equinox; the $Y$ axis points 90 degrees to the East of the $X$ axis, making the $XY$ plane the plane of the equator; The $Z$ axis points towards the North Pole. Right ascension $\alpha$ and declination $\delta$ are shown, in addition to the proper motions in their respective directions at the position of the measured galaxy.}
\label{fig:eqsys}
\end{figure}

\par The heliocentric galactic coordinate system (often ambiguously referred to as galactic coordinates), is a different spherical coordinate system centered at the Sun.  The heliocentric fundamental plane is aligned with the galactic plane, and the primary direction is along the line connecting the sun and the galactic center.  Positions are measured in galactic latitude and longitude. Latitude, $b$, measures angular distance of an object perpendicular to the galactic plane, with positive angles to the north galactic pole. Longitude, $l$, is the angular distance measured eastward along the galactic equator from the galactic center.

\begin{figure}[ht]
\includegraphics[width=\linewidth]{figures/hel_sys.png}
\caption{Illustration of the heliocentric coordinate system. The origin is the center of the Sun. The $X$ axis points towards the center of the MW; the $Y$ axis points in the direction of the sun's rotation about the galactic center; The $Z$ axis points towards the galactic North Pole. Galactic longitude $l$ and latitude $b$ are shown, in addition to the proper motions in their respective directions at the position of the measured galaxy. $\hat{v}_{sys}$ shows the unit vector in the direction of systemic line of sight velocity for the observed galaxy.}
\label{fig:hel_sys}
\end{figure}

\par The galactocentric coordinate system used in this paper is the same as the one used by van der Marel \citep{VdM08}.  The $Z$ axis points towards the galactic north pole, the $Y$ axis is oriented in the direction of the sun's rotation about the galactic center, and the $X$ axis points from the sun towards the galactic center. This is a standard, right-handed coordinate system. The galactocentric cartesian coordinate system differs from the heliocentric cartesian coordinate system only in the choice of origin, so the transformation between them is a simple translation.

\subsection{Equatorial to Galactic Transformation}
\label{sec:eq_to_gal}
\par 
We require two transformations from Equatorial to Galactic coordinates: one for the positions $(\alpha, \delta)\rightarrow (l,b)$ and one for the proper motions $(\mu_{\alpha*}, \mu_{\delta})\rightarrow (\mu_{l*}, \mu_{b})$. For both transformations, we follow the method described by \cite{poleski2013}, solving the following system of equations to find the Galactic coordinates:
\begin{equation}
\begin{aligned}
\sin b &= \cos\delta\cos\delta_{G}\cos(\alpha - \alpha_{G}) + \sin\delta\sin\delta_{G} \\
\sin(l_{\Omega} - l)\cos b &= \cos\delta\sin(\alpha - \alpha_{G}) \\
\cos(l_{\Omega} - l)\cos b &= \sin\delta\cos\delta_{G} - \cos\delta\sin\delta_{G}\cos(\alpha - \alpha_{G})
\end{aligned}
\end{equation}

where $\alpha_{G} = 192.85948$ and $\delta_{G} = 27.12825$ are the equatorial coordinates of the galactic north pole, and $l_{\Omega} = 32.93192$ is the galactic longitude of the ascending node of the galactic plane.
\par
From Poleski, the transformation for the proper motion is:
\begin{equation}
\left( \begin{array}{c}
\mu_{l*} \\
\mu_{b} \\
\end{array} \right) = 
\frac{1}{\cos b}
\left( \begin{array}{cc}
C_{1} & C_{2} \\
-C_{2} & C_{1} \\
\end{array} \right)
\left( \begin{array}{c}
\mu_{\alpha*} \\
\mu_{\delta} \\
\end{array} \right)
\end{equation}
where...
\begin{equation}
\begin{aligned}
C_{1}&=\sin\delta_{G}\cos\delta - \cos\delta_{G}\sin\delta\cos(\alpha-\alpha_{G})\\
C_{2}&=\cos\delta_{G}\sin(\alpha-\alpha_{G})\\
\cos b&=\sqrt[]{C_{1}^2+C_{2}^2}
\end{aligned}
\end{equation}

\subsection{Galactic Spherical to Galactocentric Cartesian}
\label{sec:sphere_to_cart}

As shown in \Fref{fig:hel_sys}, the vectors \textbf{$\hat{v}_{sys}$, $\hat{v}_{l*}$, $\hat{v}_{b}$} form an orthonormal basis for the heliocentric cartesian coordinate system. 
Given the Galactic coordinates $(l,b)$ of an observed galaxy, we can find the orthonormal basis vectors:
\begin{equation}
\begin{aligned}
\mathbf{\hat{v}}_{sys}&=(\cos l\cos b, \sin l\cos b, \sin b)\\
\mathbf{\hat{v}}_{l*}&=\mathbf{\hat{v}}_{sys}\times(-\hat{Z})\\
\mathbf{\hat{v}}_{b}&=\mathbf{\hat{v}}_{sys}\times\mathbf{\hat{v}}_{l*}
\end{aligned}
\end{equation}
The unit vector in the line of sight direction is easily found using trigonometry. The $\hat{v}_{l*}$ unit vector has no $\hat{Z}$ component and must be orthogonal to $\hat{v}_{sys}$. Knowing these two unit vectors determines the value of $\hat{v}_{b}$.
Given this basis, we can use some simple linear algebra to find the Cartesian velocity components. 
Let $\mathbf{v}=(v_{X}, v_{Y}, v_{Z})$ and $\mathbf{\tilde{v}}=(v_{sys}, v_{l*}, v_{b})$. 
Let $M$ be the change of basis matrix with columns equal to the basis vectors above. 
Applying the change of basis formula, we have:
\begin{equation}
\begin{aligned}
\mathbf{\tilde{v}}&=M^{-1}\mathbf{v}\\
M\mathbf{\tilde{v}}&=\mathbf{v}\\
\mathbf{\tilde{v}}^{T}M^{T}&=\mathbf{v}^{T}
\end{aligned}
\end{equation}

The final steps are to correct the velocities for the reflex motion of the sun and to find the galactocentric position with a simple translation. 
We follow the method in \citep{van2002new}, but use the updated values for the position and velocity vectors of the sun.
\begin{equation}
\begin{aligned}
\mathbf{r}_{\odot} &= (-R_{0},0,0)\\
\mathbf{v}_{\odot} &= (U_{\odot}, V_{0}+V_{\odot}, W_{\odot})
\end{aligned}
\end{equation}

where $R_{0} = 8.29\pm 0.16$ kpc and $V_{0} = 239 \pm 5$ km/s are from 
\cite{mcmillan2011mass} 
and the peculiar velocity of the sun with respect to the Local Standard of Rest is $(U_{\odot}, V_{\odot}, W_{\odot}) = (11.1, 12.24, 7.25)$ from 
%\cite{schonrich2010local}. 
The Galactocentric Cartesian position and velocity vectors of the observed galaxy are then:
\begin{equation}
\begin{aligned}
\mathbf{r} &= D\mathbf{\hat{v}}_{sys} + \mathbf{r}_{\odot}\\
\mathbf{v} &= \mathbf{\tilde{v}}^{T}M^{T} + \mathbf{v}_{\odot}
\end{aligned}
\end{equation}

\subsection{MW Galactocentric Observations}
\label{sec:coord_check}
We calculate the Milky Way Galactocentric Cartesian position and velocity vectors of M31, M33, and the LMC and compare to van der Marel's results as a sanity check of our coordinate transformations. 
%\Tref{table:properties} 
shows the results of our calculations. 
Our positions and velocities match those of van der Marel almost exactly.

\def\arraystretch{1.4}
\capstartfalse
\begin{deluxetable}{cccc}[h]
\tabletypesize{\footnotesize}
\tablecaption{\label{table:properties}
\tablewidth{250pt}
MW Galactocentric kinematic properties of M31, M33, LMC.}
\startdata
\hline
\hline
                      & M31               & M33               & LMC          \\
\hline
$\rx$ / Mpc  & $-0.378 \pm 0.019$ & $-0.476 \pm 0.014$ & $-0.001 \pm 0.001$   \\
$\ry$ / Mpc  & $ 0.613 \pm 0.032$ & $ 0.491 \pm 0.014$ & $-0.041 \pm 0.008$   \\
$\rz$ / Mpc  & $-0.283 \pm 0.015$ & $-0.413 \pm 0.012$ & $0.028 \pm 0.006$   \\
$\vx$ / km s$^{-1}$   & $  66.213 \pm 26.893$ & $  43.055 \pm 21.313$ & $-56.519 \pm ???$   \\
$\vy$ / km s$^{-1}$   & $ -76.264 \pm 19.079$ & $ 101.326 \pm 23.424$ & $-224.889 \pm ???$   \\
$\vz$ / km s$^{-1}$   & $  45.008 \pm 26.463$ & $ 138.894 \pm 28.114$ & $220.139 \pm ???$   \\
\hline

\enddata
\tablecomments{
$(\rx,\ry,\rz)$ is a position vector relative to the Galactic center, while
$(\vx,\vy,\vz)$ is a three-dimensional Galactocentric velocity vector. The
components of these vectors were used to derive distances and  radial and
tangential velocities, which are the inputs to the
likelihood calculations in the text. Uncertainties on quantities 
in this table were calculated by Monte
Carlo. For comparison to van der Marel see \citep{vdm12m31} for M31 and M33 and ??? for LMC.
%
%References are:
%vdM02~=~\citet{vanderMarel12}
}
\medskip
\end{deluxetable}
\capstarttrue

\subsection{Note on West and North}
\label{sec:WN}
In the literature, it is common to measure proper motion of galaxies in the North-West coordinate system. This can be confusing because often times authors do not define these directions in the text, but rather in footnotes. Most often, West is the opposite direction of right ascension, $(\mu_{W}=-\mu_{\alpha})$, and North is the same direction as declination.

\section{GMM's}
\label{sec:gmm_gof_app}

In this section, we present a detailed description of the GMM's used in this work and analyze how well they approximate their underlying distributions.

\subsection{Approximating the Likelihood}
\label{sec:lhood_app}

We use the observables in Table \ref{table:obs} to fit the likelihood function, \LikePR. For the rest of this paper, an ``observed'' Local Group is defined to be a LG system (either Quad or Pair) whose observational properties are consistent with those listed in Table \ref{table:obs}. The properties in Table \ref{table:obs} are assumed to be Gaussian with mean given by the tabular value, and standard deviation given by the uncertainty. For a given ``observed'' LG, its ``observed'' values are drawn from the corresponding Gaussian. In order to fit the likelihood function, we use the following steps:

\begin{enumerate}
\item Draw a large sample of ``observed'' Local Groups.
\item Transform the ``observed'' properties to the relevant (MW or M31) galactocentric frame.
\item Calculate the likelihood parameters, $(D,v_{r},v_{t})$
\item Fit the resulting nine (three) dimensional probability distribution for the ``observed'' Quads (Pairs).
\end{enumerate}

A large ``observed'' LG sample thus gives us the distribution of the likelihood parameters, $(D,v_{r},v_{t})$ for systems similar to the Local Group. In other words, this distribution is $Pr(d\vert x)$ as a function of $x$, where $x=(D,v_{r},v_{t},M_{\text{vir}},C_{\text{vir}},...)$ are the properties of a \consuelo halo and $d$ is the observational data from Table \ref{table:obs}. We fit the above distribution with a Gaussian Mixture Model (GMM) in order to evaluate likelihoods for \consuelo LG analog systems. The GMM has a large number of components because we are only interested in accurately reproducing the probability density function. Thus overfitting is not an issue.

\subsection{Approximating the Prior}
\label{sec:gmm_prior}

Given the isolation criteria, a sample of LG analogs can be found in the \consuelo catalogs.  However, the effective number of samples is determined by the weights assigned to the \consuelo LG analogs.

In this work, we evaluate the effective number of samples, $N95$, by finding the number of LG analogs that contribute the most to 95 percent of the cumulative likelihood. 
Thus, a large unweighted sample size can effectively only contribute a small number of LG analogs that most resemble the true Local Group. 
This issue is exacerbated by large numbers of constraints in the likelihood function.

For Quads, an entire \consuelo box contributes only one or two effective LG analogs which is too few to establish tight posterior mass estimates.  
We therefore require a method that can drastically increase the effective sample size without reducing the dimensionality of our likelihood function. 
This is achieved by modeling the \consuelo prior with a high dimensional Gaussian mixture model. The PGMM allows arbitrarily large fake halo catalogs to be drawn. 
\Sref{sec:gmm_gof} discusses how well the GMM's used in this work fit their underlying distributions. A good fit is particularly important for the PGMM.

\subsection{GMM Accuracy}
\label{sec:gmm_gof}

The accuracy of our posterior mass predictions strongly depends on how well the GMM's approximate the respective distributions. Figure \ref{fig:LGMM} shows the LGMM fit to the observed data for Quads. 
The $1\sigma$, $2\sigma$, and $3\sigma$ contours for the LGMM distribution almost perfectly overlap the corresponding contours for the observed data.  Therefore we can confidently conclude that our likelihood model is extremely accurate.

\begin{figure*}
\includegraphics[width=\linewidth]{figures/LGMM.pdf}
\caption{1D and 2D marginalizations of the likelihood function for Quads. The Red contours are generated by the LGMM model and almost perfectly approximate the true distribution.}
\label{fig:LGMM}
\end{figure*}

The second GMM used in this work is the PGMM which models the \consuelo prior. It is particularly important for the model to fit the data extremely well in this case, because we sample directly from the PGMM distribution instead of the true \consuelo prior. Figure \ref{fig:Q_GMMP_GOF} shows the goodness of fit of the PGMM for prior marginalizations. 
As with the LGMM, the $1\sigma$, $2\sigma$, and $3\sigma$ contours almost perfectly overlap. 
This is particularly encouraging for combinations of parameters that are completely independent, like $D^{LMC}$ and $D^{M31}$. 
In this paper, the PGMM has 30 components, and we verify that the posterior estimates do not vary with small changes in the number of components. 
For relatively low dimensional data sets like the ones in this paper, the 1D and 2D marginalizations provide compelling evidence that the PGMM is accurately modeling the \consuelo prior. 
However for high dimensional data sets, the 1D and 2D marginalizations may not be a sufficient goodness of fit test.
In this case, a generalized chi-squared test may be necessary.
Fitting the prior PDF for halo characteristics is a new approach that has not been explored in the literature. 
This method is extremely important because it allows us to create arbitrarily large fake halo catalogs. 
These fake catalogs can contain hundreds of billions of halos, a sample size that is impossible to achieve with current computation limits.
\begin{figure*}
\includegraphics[width=\linewidth]{figures/Q_GMMP_GOF.pdf}
\caption{1D and 2D marginalizations of the Consuelo prior, approximated by the PGMM. The true \consuelo contours (cyan) overlap well with the PGMM contours (magenta).}
\label{fig:Q_GMMP_GOF}
\end{figure*}

\end{document}
